### 2025-05-30 article.tex
Add initial JCGT form

\documentclass{jcgt}

\setciteauthor{Matthias Raab and Laurent Belcour and Frankie Liu and Jamie Portsmouth}
\setcitetitle{The Minimal Retroreflective Microfacet Model (MRMM)}
%\setheadtitle{Abbreviated title, only if full title won't fit in page headers}

% Mark submissions with the date of submission using the following line:
\submitted{\today}

% Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
\accepted{yyyy-mm-dd}{yyyy-mm-dd}{yyyy-mm-dd}{Editor Name}{volume}{issue}{1}{6}{yyyy}
\seturl{http://jcgt.org/published/vol/issue/num/}

\usepackage{amsmath}

\newcommand{\vecv}{\mathbf{v}}
\newcommand{\vecvv}{\mathbf{v'}}
\newcommand{\vecl}{\mathbf{l}}
\newcommand{\vecll}{\mathbf{l'}}
\newcommand{\vecn}{\mathbf{n}}
\newcommand{\vech}{\mathbf{h}}
\newcommand{\vecb}{\mathbf{b}}

\newcommand{\vdot}[2]{\left\langle #1, #2 \right\rangle}

\begin{document}

\author{Matthias Raab\\NVIDIA
        \and Laurent Belcour\\Intel
        \and Frankie Liu\\NVIDIA
        \and Jamie Portsmouth\\Autodesk
       }

\title{The Minimal Retroreflective Microfacet Model}

\maketitle
\thispagestyle{firstpagestyle}

\begin{abstract}
  \small
  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for
  modifying a pre-existing microfacet BRDF implementation to produce a visual plausible retroreflective result.
  \end{abstract}

\section{Introduction}

\section{Microfacet BRDF Models}

\begin{figure}
  \begin{center}
    \begin{tabular} {|l|l|}
      \hline
      $\vecv$ & normalized vector to viewer \\
      \hline
      $\vecl$ & normalized vector to light \\
      \hline
      $\vecn$ & surface normal \\
      \hline
      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent  \\
      \hline
      $\vech$ & half vector, $\vech = \vech(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
      \hline
      $\vdot{\mathbf{x}}{\mathbf{y}}$ & dot product of $\mathbf{x}$ and $\mathbf{y}$ \\
      \hline
      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $-\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
      \hline
      $\Omega$ & hemisphere of directions around the normal \\
      \hline
    \end{tabular}
  \end{center}
  \caption{Notations used.}
  \label{fig:notation}
\end{figure}

Using the notations from figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form
\begin{equation}\label{eq:microfacet}
  f(\vecv, \vecl) = \frac{D(\vech) G_{2}(\vecv,\vecl)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
\end{equation}
where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$,
and $G_{2}$ is the shadowing-masking function.
The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
$G_1(\vecl, \vecv)$ for light direction:
\begin{equation}
G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right)
\end{equation}
Mathematically, the BRDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
\begin{equation}\label{eq:visible}
  \vdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
\end{equation}
Choices for $G_1$ with this property are the Smith masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\vdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
\end{equation}
where analytic expressions are available for Beckmann and GGX distributions, and the v-cavity masking function
\begin{equation}
  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
  \min\left\lbrace\frac{2\vdot{\vecv}{\vecn}\vdot{\vecn}{\vech}}{\vdot{\vecv}{\vech}}, 1\right\rbrace,
\end{equation}
which is generally applicable for distributions with symmetry around the normal.

\section{Back-vector Modification}

If instead of half-vector $\vech$ the back-vector $\vecb$, defined as
\begin{equation}
  \vecb(\vecv, \vecl) = \vech(\vecvv, \vecl) = \frac{\vecvv + \vecl}{\lVert \vecvv + \vecl \rVert}, \text{ with } \vecvv = \text{reflect}(\vecv, \vecn),
\end{equation}
is used in equation \eqref{eq:microfacet}, the resulting BRDF is retro-reflective \cite{BelcourBackvector}.
Note that instead of explicitly defining a BRDF using the back-vector, we can simply use $\vecvv$ instead of $\vecv$, i.e. we define $f_{\text{retro}}(\vecv, \vecl) := f(\vecvv, \vecl)$ such that the half-vector becomes the back-vector.
In the following we show that this BRDF is plausible, i.e. symmetry and energy conservation are guaranteed.

\section{Plausibility of the Back-vector Modification}

In analogy to $\vecvv$ we let $\vecll := \text{reflect}(\vecl, \vecn)$.
First we observe
\begin{equation}\label{eq:nk}
  \begin{split}
    \vdot{\vecvv}{\vecn} &= \vdot{\vecv}{\vecn} \text{ and}\\
    \vdot{\vecl}{\vecn} &= \vdot{\vecll}{\vecn}.
  \end{split}
\end{equation}
For the back-vector we show $\vech(\vecvv, \vecl) = \text{reflect}\left(\vech(\vecll, \vecv), \vecn\right)$, as
\begin{equation}
  \begin{split}
  \vecvv + \vecl &=
  \left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \vecl \\
  &= \left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \left(- \vecll + 2 \vdot{\vecll}{\vecn} \vecn \right)\\
  &= -(\vecll + \vecv) + 2 \vdot{(\vecll + \vecv)}{\vecn} \vecn =
  \text{reflect}(\vecll + \vecv, \vecn)
  \end{split}
\end{equation}
and therefore $\lVert \vecvv + \vecl \rVert = \lVert \vecll + \vecv \rVert$ as well.
From that we can conclude
\begin{equation}\label{eq:nh}
  \begin{split}
    \vdot{\vecn}{\vech(\vecvv, \vecl)} &=
    \vdot{\vecn}{\vech(\vecll, \vecv)}\\
    \vdot{\mathbf{t_1}}{\vech(\vecvv, \vecl)} &=
    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)}\\
    \vdot{\mathbf{t_2}}{\vech(\vecvv, \vecl)} &=
    -\vdot{\mathbf{t_2}}{\vech(\vecll, \vecv)}.\\
    \end{split}
\end{equation}
As anisotropic Phong, Beckmann, and GGX distributions are reflection-symmetric around the normal, i.e. they take the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using them with the back-vector obeys symmetry.
Further, the reflection symmetry of $D$ along with equation \eqref{eq:nk} gives
\begin{equation}
  \begin{split}
  G_{1,\text{Smith}}(\vecv) &=
  \frac{\vdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
  &= \frac{\vdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =
  G_{1,\text{Smith}}(\vecvv)
  \end{split}
\end{equation}
for Smith-masking. Next we show
\begin{equation}
  \vdot{\vecvv}{\vecl} = -\vdot{\vecv}{\vecl} + 2 \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn}
  = \vdot{\vecv}{-\vecl + 2 \vdot{\vecl}{\vecn}\vecn} = \vdot{\vecv}{\vecll}
\end{equation}
and therefore
\begin{equation}\label{eq:kh}
  \vdot{\vecvv}{\vech(\vecvv, \vecl)} =
  \frac{1 + \vdot{\vecl}{\vecvv}}{\lVert \vecvv + \vecl \rVert} =
  \frac{1 + \vdot{\vecll}{\vecv}}{\lVert \vecll + \vecv \rVert} =
  \vdot{\vecll}{\vech(\vecll,\vecv)}.
\end{equation}
Using this along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
\begin{equation}
  \begin{split}
  G_1(\vecvv, \vecl) &=
  G_{1, \text{vc}}\left( \vdot{\vecvv}{\vecn}, \vdot{\vecn}{\vech(\vecvv, \vecl)}, \vdot{\vecvv}{\vech(\vecvv, \vecl)} \right)\\
  &=  G_{1, \text{vc}}\left( \vdot{\vecv}{\vecn}, \vdot{\vecn}{\vech(\vecll, \vecv)}, \vdot{\vecll}{\vech(\vecll, \vecv)} \right) \\
  &=  G_{1, \text{vc}}\left( \vdot{\vecv}{\vecn}, \vdot{\vecn}{\vech(\vecv, \vecll)}, \vdot{\vecll}{\vech(\vecv, \vecll)} \right) \\
  &= G_1(\vecv, \vecll)
  \end{split}
\end{equation}
and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the v-cavity masking function.
In summary, we can conclude the symmetry of the BRDF
\begin{equation}
  \begin{split}
  f_{\text{retro}}(\vecv, \vecl) =
  f(\vecvv, \vecl) &=
  \frac{D(\vech(\vecvv, \vecl)) G_{2}(G_1(\vecvv, \vecl), G_1(\vecl, \vecvv))}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
  &=
  \frac{D(\vech(\vecll, \vecv)) G_{2}(G_1(\vecv, \vecll), G_1(\vecll, \vecv))}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} } =
  f(\vecll, \vecv) =
  f_{\text{retro}}(\vecl, \vecv).
  \end{split}
\end{equation}
For the albedo we simply have
\begin{equation}
  \rho_{\text{retro}}(\vecv) =
  \int_\Omega f_{\text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
  \int_\Omega f(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho(\vecvv)
\end{equation}
i.e. we get the albedo of the regular microfacet BRDF for the reflected direction.
Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
\begin{equation}
  \begin{split}
  \rho(\vecvv) &=
  \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
  &\leq
  \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
  = 1.
  \end{split}
\end{equation}

\section{Implementation Notes}

Given an implementation of a regular microfacet BRDF, extending it to retro-reflection is straightforward:

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BRDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BRDF and retro-reflective BRDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
\end{itemize}

\bibliographystyle{jcgt}
\bibliography{citations}

\section*{Author Contact Information}

\hspace{-2mm}\begin{tabular}{p{0.48\textwidth}p{0.48\textwidth}}
Matthias Raab \newline
\href{mailto:matthias.raab@gmail.com}{matthias.raab@gmail.com} \newline
&
Laurent Belcour \newline
\href{mailto:laurent.belcour@gmail.com}{laurent.belcour@gmail.com} \\
Frankie Liu \newline
\href{mailto:frankie.liu@gmail.com}{frankie.liu@gmail.com} \newline
&
Jamie Portsmouth \newline
\href{mailto:jamports@mac.com}{jamports@mac.com} \newline
\end{tabular}

\end{document}

### 2025-06-26 article.tex
Merging Laurent's work. Adding some images, and tweaks.

@@
 \documentclass{jcgt}
+
+\usepackage{amsmath}
+\usepackage{listings}
+\usepackage{xcolor}
+\usepackage[normalem]{ulem}
+\usepackage{siunitx}
+\usepackage{csquotes}
+\usepackage{url}
+\usepackage{subcaption}
+\usepackage{xcolor}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{mathtools}
+\usepackage{xfrac}
+\usepackage{comment}
+\usepackage{natbib}
 
@@
 
-\usepackage{amsmath}
-
 \newcommand{\vecv}{\mathbf{v}}
@@
 \newcommand{\vecn}{\mathbf{n}}
+\newcommand{\vecm}{\mathbf{m}}
 \newcommand{\vech}{\mathbf{h}}
 \newcommand{\vecb}{\mathbf{b}}
-
 \newcommand{\vdot}[2]{\left\langle #1, #2 \right\rangle}
 
+\newcommand{\todo}[1]{{\color{red}\textbf{[TODO:} #1\textbf{]}}}
+\newcommand{\laurent}[1]{{\color{blue}\textbf{[Laurent:} #1\textbf{]}}}
+
+\includecomment{introduction}
+\includecomment{existing-models}
+\includecomment{eon}
+\includecomment{eon-sampling}
+\includecomment{conclusion}
+
+\urlstyle{same}
+
+\captionsetup[table]{skip=10pt}
+
+\lstdefinelanguage{GLSL}%
+{%
+	morekeywords={%
+	% HLSL constants
+		false,FALSE,NULL,true,TRUE,%
+	% GLSL predefinde macro constant
+		__LINE__,__FILE__,__VERSION__,GL_core_profile,GL_es_profile,GL_compatibility_profile,%
+	% GLSL precision modifier
+		precision,highp,mediump,lowp,%
+	% GLSL types
+		void,bool,int,uint,float,double,vec2,vec3,vec4,dvec2,dvec3,dvec4,bvec2,bvec3,bvec4,ivec2,ivec3,ivec4,uvec2,uvec3,uvec4,mat2,mat3,mat4,mat2x2,mat2x3,mat2x4,mat3x2,mat3x3,mat3x4,mat4x2,mat4x3,mat4x4,dmat2,dmat3,dmat4,dmat2x2,dmat2x3,dmat2x4,dmat3x2,dmat3x3,dmat3x4,dmat4x2,dmat4x3,dmat4x4,sampler1D,sampler2D,sampler3D,image1D,image2D,image3D,samplerCube,imageCube,sampler2DRect,image2DRect,sampler1DArray,sampler2DArray,image1DArray,image2DArray,samplerBuffer,imageBuffer,sampler2DMS,image2DMS,sampler2DMSArray,image2DMSArray,samplerCubeArray,imageCubeArray,sampler1DShadow,sampler2DShadow,sampler2DRectShadow,sampler1DArrayShadow,sampler2DArrayShadow,samplerCubeShadow,samplerCubeArrayShadow,isampler1D,isampler2D,isampler3D,iimage1D,iimage2D,iimage3D,isamplerCube,iimageCube,isampler2DRect,iimage2DRect,isampler1DArray,isampler2DArray,iimage1DArray,iimage2DArray,isamplerBuffer,iimageBuffer,isampler2DMS,iimage2DMS,isampler2DMSArray,iimage2DMSArray,isamplerCubeArray,iimageCubeArray,atomic_uint,usampler1D,usampler2D,usampler3D,uimage1D,uimage2D,uimage3D,usamplerCube,uimageCube,usampler2DRect,uimage2DRect,usampler1DArray,usampler2DArray,uimage1DArray,uimage2DArray,usamplerBuffer,uimageBuffer,usampler2DMS,uimage2DMS,usampler2DMSArray,uimage2DMSArray,usamplerCubeArray,uimageCubeArray,struct,%
+	% GLSL support variables
+		gl_BackColor,gl_BackLightModelProduct,gl_BackLightProduct,gl_BackMaterial,gl_BackSecondaryColor,gl_ClipDistance,gl_ClipPlane,gl_ClipVertex,gl_Color,gl_DepthRange,gl_DepthRangeParameters,gl_EyePlaneQ,gl_EyePlaneR,gl_EyePlaneS,gl_EyePlaneT,gl_Fog,gl_FogCoord,gl_FogFragCoord,gl_FogParameters,gl_FragColor,gl_FragCoord,gl_FragData,gl_FragDepth,gl_FrontColor,gl_FrontFacing,gl_FrontLightModelProduct,gl_FrontLightProduct,gl_FrontMaterial,gl_FrontSecondaryColor,gl_InstanceID,gl_Layer,gl_LightModel,gl_LightModelParameters,gl_LightModelProducts,gl_LightProducts,gl_LightSource,gl_LightSourceParameters,gl_MaterialParameters,gl_ModelViewMatrix,gl_ModelViewMatrixInverse,gl_ModelViewMatrixInverseTranspose,gl_ModelViewMatrixTranspose,gl_ModelViewProjectionMatrix,gl_ModelViewProjectionMatrixInverse,gl_ModelViewProjectionMatrixInverseTranspose,gl_ModelViewProjectionMatrixTranspose,gl_MultiTexCoord0,gl_MultiTexCoord1,gl_MultiTexCoord2,gl_MultiTexCoord3,gl_MultiTexCoord4,gl_MultiTexCoord5,gl_MultiTexCoord6,gl_MultiTexCoord7,gl_Normal,gl_NormalMatrix,gl_NormalScale,gl_ObjectPlaneQ,gl_ObjectPlaneR,gl_ObjectPlaneS,gl_ObjectPlaneT,gl_Point,gl_PointCoord,gl_PointParameters,gl_PointSize,gl_Position,gl_PrimitiveIDIn,gl_ProjectionMatrix,gl_ProjectionMatrixInverse,gl_ProjectionMatrixInverseTranspose,gl_ProjectionMatrixTranspose,gl_SecondaryColor,gl_TexCoord,gl_TextureEnvColor,gl_TextureMatrix,gl_TextureMatrixInverse,gl_TextureMatrixInverseTranspose,gl_TextureMatrixTranspose,gl_Vertex,gl_VertexID,%
+	% GLSL support constants
+		gl_MaxClipPlanes,gl_MaxCombinedTextureImageUnits,gl_MaxDrawBuffers,gl_MaxFragmentUniformComponents,gl_MaxLights,gl_MaxTextureCoords,gl_MaxTextureImageUnits,gl_MaxTextureUnits,gl_MaxVaryingFloats,gl_MaxVertexAttribs,gl_MaxVertexTextureImageUnits,gl_MaxVertexUniformComponents,%
+	% GLSL support functions
+		abs,acos,all,any,asin,atan,ceil,clamp,cos,cross,degrees,dFdx,dFdy,distance,dot,equal,exp,exp2,faceforward,floor,fract,ftransform,fwidth,greaterThan,greaterThanEqual,inversesqrt,length,lessThan,lessThanEqual,log,log2,matrixCompMult,max,min,mix,mod,noise1,noise2,noise3,noise4,normalize,not,notEqual,outerProduct,pow,radians,reflect,refract,shadow1D,shadow1DLod,shadow1DProj,shadow1DProjLod,shadow2D,shadow2DLod,shadow2DProj,shadow2DProjLod,sign,sin,smoothstep,sqrt,step,tan,texture1D,texture1DLod,texture1DProj,texture1DProjLod,texture2D,texture2DLod,texture2DProj,texture2DProjLod,texture3D,texture3DLod,texture3DProj,texture3DProjLod,textureCube,textureCubeLod,transpose,%
+	% GLSL struct member -> FixMe: Should have dot(.) as delimiter
+		rgb
+	},
+	sensitive=true,%
+	morecomment=[s]{/*}{*/},%
+	morecomment=[l]//,%
+	morestring=[b]",%
+	morestring=[b]',%
+	moredelim=*[directive]\#,%
+	% keyword.control.hlsl
+	moredirectives={define,defined,elif,else,if,ifdef,endif,line,error,ifndef,include,pragma,undef,warning,extension,version},%
+  % keyword highlighting%
+  classoffset=1, % starting new class
+  keywords={retroGGX, break,case,continue,default,discard,do,else,for,if,return,switch,while,define},%
+  keywordstyle=\color{codeblue}\bfseries,%
+  classoffset=0%
+  }[keywords,comments,strings,directives]%
+
+\definecolor{backcolour}{rgb}{1, 1, 1}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{codeblue}{rgb}{0,0.3,0.6}
+
+\lstset{language=GLSL,
+        linewidth=1.0\linewidth,
+        backgroundcolor=\color{backcolour},
+        commentstyle=\color{codegreen},
+        keywordstyle=\color{magenta},
+        numberstyle=\tiny\color{codegray},
+        stringstyle=\color{codepurple},
+        basicstyle=\ttfamily\scriptsize,
+        breakatwhitespace=false,
+        breaklines=true,
+        keepspaces=true,
+        numbers=none,
+        numbersep=5pt,
+        showspaces=false,
+        showstringspaces=false,
+        showtabs=false,
+        escapeinside={<@}{@>},
+        tabsize=2,
+        texcl=false,
+        frameround=fttt
+        }
+
 \begin{document}
+
+\title{The Minimal Retroreflective Microfacet Model}
 
@@
 
-\title{The Minimal Retroreflective Microfacet Model}
+\teaser{
+  \vspace*{-0.5cm}
+  \centering
+    \includegraphics[width=0.32\linewidth]{figures/sphere_side_r0.3.png     }
+    \includegraphics[width=0.32\linewidth]{figures/sphere_quarter_r0.3.png  }
+    \includegraphics[width=0.32\linewidth]{figures/sphere_front_r0.3.png    }
+  \caption{Comparison of our retroreflector model (top) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (bottom).
+The lighting varies (left to right) from side-on, quarter-on, and straight-on, exhibiting the strong retro-reflective peak.
+   \vspace{-0.5cm}
+  \label{fig:teaser}}
+}
 
@@
   \small
-  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for
-  modifying a pre-existing microfacet BRDF implementation to produce a visual plausible retroreflective result.
+  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BRDF implementation to produce a visually plausible, energy-conserving retroreflective result.
   \end{abstract}
@@
 
+Rendering of retro-reflective materials is useful in a variety of contexts.
+A number of CG models have been proposed.
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
+We present here a model based on a the previously published back-vector formulation~\cite{BelcourBackvector} which meets are these requirements.
+It is so simple a modification to a regular microfacet model, that we `Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
+
+In this document, we detail the correct usage and derivation of this retro-reflective BRDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
+We first recall some basic principles from microfacets theory (Section~\ref{sec:microfacets}) and show how to change it to account for retro-reflective microfacets (Section~\ref{sec:backvector}).
+We then detail how to implement this in practice (Section~\ref{sec:implementation}).
+Finally, we discuss how retro-reflection can occur at the microfacet level (Section~\ref{sec:plausibility}).
+
 \section{Microfacet BRDF Models}
+\label{sec:microfacets}
 
@@
       \hline
-      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent  \\
+      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
       \hline
@@
       \hline
-      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e. $-\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
+      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e.
+$\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
       \hline
@@
 
-Using the notations from figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form
-\begin{equation}\label{eq:microfacet}
-  f(\vecv, \vecl) = \frac{D(\vech) G_{2}(\vecv,\vecl)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
+Using the notations from figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
+\begin{equation}
+  f(\vecv, \vecl) = \frac{D(\vech) \, G_{2}(\vecv,\vecl) \, F(\vech, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
+  \label{eq:microfacet_model}
 \end{equation}
-where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$,
-and $G_{2}$ is the shadowing-masking function.
+where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
+The half-vector $\vech$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacet with normals aligned with the half-vector will reflect light).
+In Section~4.1 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equation~\ref{eq:microfacet_model}.
+
+In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector.
+Before detailing the impact of such a change, we review some properties of microfacet models.
+
+\paragraph*{Shadowing-Masking.}
 The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
@@
 \section{Back-vector Modification}
+\label{sec:backvector}
 
-If instead of half-vector $\vech$ the back-vector $\vecb$, defined as
+If instead of half-vector $\vech$, we use the back-vector $\vecb$~\cite{BelcourBackvector} defined as
 \begin{equation}
@@
 \end{equation}
-is used in equation \eqref{eq:microfacet}, the resulting BRDF is retro-reflective \cite{BelcourBackvector}.
-Note that instead of explicitly defining a BRDF using the back-vector, we can simply use $\vecvv$ instead of $\vecv$, i.e.
-we define $f_{\text{retro}}(\vecv, \vecl) := f(\vecvv, \vecl)$ such that the half-vector becomes the back-vector.
-In the following we show that this BRDF is plausible, i.e. symmetry and energy conservation are guaranteed.
+to select the reflecting microfacet, we are to re-derive the microfacet model.
+
+\paragraph*{Jacobian of Back-Vector.}
+The first step is to write the Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}).
+Fortunately, this jacobian has the same form as the half-vector one:
+\begin{equation}
+\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vecl \cdot \vech(\vecvv, \vecl) \right|} = \dfrac{1}{4 \left| \vecl \cdot \vecb \right|}
+\end{equation}
+
+\paragraph*{Microfacet reflection term.}
+We do not know for sure how the retro-reflective microfacet interacts with incident light.
+However, we have experimental evidence~\cite{BelcourBackvector} that it behaves similarly to a Fresnel term.
+We opted to reuse the Fresnel as well in the $(\vecvv, \vecl)$ configuration.
+As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead.
+It follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
+\begin{equation}
+  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vecvv \cdot \vecb \right|^2}.
+\end{equation}
+We used the fact that $\left| \vecl \cdot \vecb \right| = \left| \vecvv \cdot \vecb \right|$.
+
+\paragraph*{Shadowing \& Masking.} \todo{Laurent: I think we could extract the shadowing from Matthias derivation here.
+The other way to justify the derivation of the full model is to highlight that Smith's shadowing/masking does not depend on the microfacet (appart from the sideness).
+Hence, whatever the microfacet selector (as long as the sideness matches), the shadowing/masking is the same.
+Since $\vecv$ and $\vecvv$ share the same slope, we can exchange them in the equation.}
+
+\paragraph*{Final form.}
+\begin{equation}
+  f_R(\vecv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecb, \vecvv)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  \label{eq:retro_microfacet_model}
+\end{equation}
 
 \section{Plausibility of the Back-vector Modification}
+
+\label{sec:plausibility}
 
@@
 \section{Implementation Notes}
+\label{sec:implementation}
+
+Given an implementation of a regular microfacet BRDF, extending it to retro-reflection is straightforward:
+
+\begin{itemize}
+\item
+  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
+\item
+  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BRDF.
+  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
+\item
+  As the albedos of standard BRDF and retro-reflective BRDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
+\item Pseudo-code is provided in Listing~\ref{lst:implementation}.
+\end{itemize}
+
+\noindent\begin{minipage}[tb]{\linewidth}
+\begin{lstlisting}[
+    language=GLSL,
+    caption=Pseudo code to implement the retro-reflective microfacet model.,
+    label=lst:implementation,
+    nolol=true,
+    frame=trBL
+    ]
+vec3 retroGGX(vec3 L, vec3 V, vec3 N, vec3 X, vec3 Y)
+{
+    vec3 Vp = -reflect(V,N); // jp - is this wrong? Why the minus sign
+    float NdotL = dot(N, L);
+    float NdotV = dot(N, Vp);
+    if (NdotL < 0 || NdotV < 0) return vec3(0);
+
+    vec3 B = normalize(L+Vp);
+    float NdotH = dot(N, B);
+    vec3 wB = vec3(dot(B,X),  dot(B,Y),  dot(B,N));
+    vec3 wL = vec3(dot(L,X),  dot(L,Y),  dot(L,N));
+    vec3 wV = vec3(dot(Vp,X), dot(Vp,Y), dot(Vp,N));
+
+    float D = D_GGX(wB, ax, ay);
+    float G = G2_GGX(wV, wL, ax, ay);
+
+    float c = dot(L, B);
+    float F = schlick(f0, c); // I don't think we need to give the form of the Fresnel
+
+    float val = Ks * D * G * F / (4 * NdotL * NdotV);
+    return vec3(val);
+}
+\end{lstlisting}
+\end{minipage}
+
+\section{Implementation Notes}
 

### 2025-06-26 scripts/retro.py

@@ -1,0 +1,153 @@
+
+####################################################################
+# Minimal Retroreflective Microfacet Model
+####################################################################
+
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+import colorsys
+
+DENOM_TOLERANCE       = 1.0e-20
+FLT_EPSILON           = 1.1920929e-7
+
+def sqr(X): return X*X
+def sqrt(X): return X**0.5
+
+def D_ggx(M, r):
+    ax = max(sqr(r), DENOM_TOLERANCE)
+    ay = max(sqr(r), DENOM_TOLERANCE)
+    Ddenom = math.pi * ax * ay * sqr(sqr(M[0]/ax) + sqr(M[1]/ay) + sqr(M[2]))
+    return 1.0 / max(Ddenom, DENOM_TOLERANCE)
+
+
+def ggx_lambda(V, r):
+    if abs(V[2]) < FLT_EPSILON:
+        return 0.0
+    return (-1.0 + sqrt(1.0 + (sqr(r*V[0]) + sqr(r*V[1]))/sqr(V[2]))) / 2.0
+
+def G2_ggx(V, L, m, r):
+    return 1.0 / (1.0 + ggx_lambda(V, r) + ggx_lambda(L, r))
+
+def normalize(v):
+    norm = np.linalg.norm(v)
+    if norm == 0:
+       return v
+    return v / norm
+
+def brdf_ggx(V, L, r):
+    H = normalize(V + L)
+    D = D_ggx(H, r)
+    G2 = G2_ggx(V, L, H, r)
+    J = 1.0 / max(4.0*abs(V[2])*abs(L[2]), DENOM_TOLERANCE)
+    return G2 * D * J
+
+def brdf_raab(V, L, r):
+    V = np.array([-V[0], -V[1], V[2]])
+    H = normalize(V + L)
+    D = D_ggx(H, r)
+    G2 = G2_ggx(V, L, H, r)
+    J = 1.0 / max(4.0*abs(V[2])*abs(L[2]), DENOM_TOLERANCE)
+    return G2 * D * J
+
+def brdf_belcour(V, L, r):
+    Vp = np.array([-V[0], -V[1], V[2]])
+    B = normalize(Vp + L)
+    D = D_ggx(B, r)
+    G2 = G2_ggx(V, L, B, r)
+    J = 1.0 / max(4.0*abs(V[2])*abs(L[2]), DENOM_TOLERANCE)
[…]

### 2025-06-26 article.tex
Tweak listing

@@
     ]
-vec3 retroGGX(vec3 L, vec3 V, vec3 N, vec3 X, vec3 Y)
+vec3 brdf_retro(vec3 L, vec3 V, vec3 N, vec3 X, vec3 Y)
 {
     vec3 Vp = -reflect(V,N); // jp - is this wrong? Why the minus sign
+
+    // The following could all be replaced with:
+    // We could give the simplified form of brdf_ggx, but not essential.
+    return brdf_ggx(L, Vp, N, X, Y);
+
+    /*
     float NdotL = dot(N, L);
@@
     return vec3(val);
+    */
 }

### 2025-06-26 article.tex
Typos

@@
   \caption{Comparison of our retroreflector model (top) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (bottom).
-The lighting varies (left to right) from side-on, quarter-on, and straight-on, exhibiting the strong retro-reflective peak.
+The lighting varies (left to right) from side-on, to quarter-on, and straight-on, exhibiting the strong retro-reflective peak.
    \vspace{-0.5cm}
@@
 For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
-We present here a model based on a the previously published back-vector formulation~\cite{BelcourBackvector} which meets are these requirements.
-It is so simple a modification to a regular microfacet model, that we `Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
+We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets are these requirements.
+It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
 
 In this document, we detail the correct usage and derivation of this retro-reflective BRDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
-We first recall some basic principles from microfacets theory (Section~\ref{sec:microfacets}) and show how to change it to account for retro-reflective microfacets (Section~\ref{sec:backvector}).
+We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}) and show how to account for retro-reflective microfacets (Section~\ref{sec:backvector}) within this framework.
 We then detail how to implement this in practice (Section~\ref{sec:implementation}).

### 2025-07-02 article.tex
Fix email and remove forumula

@@
 i.e. we get the albedo of the regular microfacet BRDF for the reflected direction.
-Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
-\begin{equation}
-  \begin{split}
-  \rho(\vecvv) &=
-  \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
-  &\leq
-  \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
-  = 1.
-  \end{split}
-\end{equation}
+As the original microfacet BRDF fulfills energy conversion, the retro-reflective modification does, too.
+%% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
+%% \begin{equation}
+%%   \begin{split}
+%%   \rho(\vecvv) &=
+%%   \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
+%%   &\leq
+%%   \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
+%%   = 1.
+%%   \end{split}
+%% \end{equation}
 
@@
 Matthias Raab \newline
-\href{mailto:matthias.raab@gmail.com}{matthias.raab@gmail.com} \newline
+\href{mailto:mraab@nvidia.com}{mraab@nvidia.com} \newline
 &

### 2025-07-09 article.tex
update text

@@
   \begin{center}
-    \begin{tabular} {|l|l|}
+    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
+%    \begin{tabular} {|l|l|}
       \hline
@@
 
-If instead of half-vector $\vech$, we use the back-vector $\vecb$~\cite{BelcourBackvector} defined as
+Instead of half-vector $\vech$, we can use the back-vector $\vecb$~\cite{BelcourBackvector} defined as
 \begin{equation}
@@
 \end{equation}
-to select the reflecting microfacet, we are to re-derive the microfacet model.
+to select the reflecting microfacet. %, we are to re-derive the microfacet model.
+As illustrated in figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
+Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
+While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.2\textwidth]{figures/sphere_front_r0.3.png} % TODO: add sketch
+  \end{center}
+  \caption{Geometry of the back vector, illustrated in two dimensions.}
+  \label{fig:backvector}
+\end{figure}
 
 \paragraph*{Jacobian of Back-Vector.}
-The first step is to write the Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}).
-Fortunately, this jacobian has the same form as the half-vector one:
+First we observe that  Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
 \begin{equation}
@@
 \paragraph*{Microfacet reflection term.}
-We do not know for sure how the retro-reflective microfacet interacts with incident light.
-However, we have experimental evidence~\cite{BelcourBackvector} that it behaves similarly to a Fresnel term.
-We opted to reuse the Fresnel as well in the $(\vecvv, \vecl)$ configuration.
+There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements.
 As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead.
@@
 
-\paragraph*{Shadowing \& Masking.} \todo{Laurent: I think we could extract the shadowing from Matthias derivation here.
+\paragraph*{Shadowing \& Masking.}
+We simply apply Smith or v-cavities masking to the mirrored microfacet distribution.
+With Smith masking, this is actually identical to a regular microfacet BRDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
+For v-cavities masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
+%todo{Laurent: I think we could extract the shadowing from Matthias derivation here.
 The other way to justify the derivation of the full model is to highlight that Smith's shadowing/masking does not depend on the microfacet (appart from the sideness).
@@
 \paragraph*{Final form.}
+The final retro-reflective microfacet model can be written as
 \begin{equation}
-  f_R(\vecv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecb, \vecvv)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
-  \label{eq:retro_microfacet_model}
+  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecb, \vecvv)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  \label{eq:retro_microfacet_model}.
 \end{equation}
+In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
 
@@
 
-\section{Implementation Notes}
-
-Given an implementation of a regular microfacet BRDF, extending it to retro-reflection is straightforward:
-
-\begin{itemize}
-\item
-  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
-\item
-  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BRDF.
-  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
-\item
-  As the albedos of standard BRDF and retro-reflective BRDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
-\end{itemize}
-
 \bibliographystyle{jcgt}

### 2025-07-10 article.tex
- add illustration - unify math - add Fresnel to plausibility - updates on text

@@
 \seturl{http://jcgt.org/published/vol/issue/num/}
-
 \newcommand{\vecv}{\mathbf{v}}
@@
 \newcommand{\vecb}{\mathbf{b}}
+\newcommand{\vecx}{\mathbf{x}}
+\newcommand{\vecy}{\mathbf{y}}
 \newcommand{\vdot}[2]{\left\langle #1, #2 \right\rangle}
@@
 In this document, we detail the correct usage and derivation of this retro-reflective BRDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
-We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}) and show how to account for retro-reflective microfacets (Section~\ref{sec:backvector}) within this framework.
-We then detail how to implement this in practice (Section~\ref{sec:implementation}).
-Finally, we discuss how retro-reflection can occur at the microfacet level (Section~\ref{sec:plausibility}).
+We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BRDF is fully plausible (Section~\ref{sec:plausibility}).
+Finally, we sketch how the model can be implemented in practice (Section~\ref{sec:implementation}).
 
@@
       \hline
-      $\vdot{\mathbf{x}}{\mathbf{y}}$ & dot product of $\mathbf{x}$ and $\mathbf{y}$ \\
+      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
       \hline
@@
 to select the reflecting microfacet. %, we are to re-derive the microfacet model.
-As illustrated in figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
-Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
-While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.
-\begin{figure}
+
+\begin{figure}[h]
   \begin{center}
-    \includegraphics[width=0.2\textwidth]{figures/sphere_front_r0.3.png} % TODO: add sketch
+    \includegraphics[width=0.4\textwidth]{figures/backvector.pdf}
   \end{center}
@@
 
+As illustrated in figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
+Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
+While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.
+
 \paragraph*{Jacobian of Back-Vector.}
@@
 \begin{equation}
-\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vecl \cdot \vech(\vecvv, \vecl) \right|} = \dfrac{1}{4 \left| \vecl \cdot \vecb \right|}
+\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
 \end{equation}
@@
 \begin{equation}
-  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vecvv \cdot \vecb \right|^2}.
+  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vdot{\vecvv}{\vecb} \right|^2}.
 \end{equation}
-We used the fact that $\left| \vecl \cdot \vecb \right| = \left| \vecvv \cdot \vecb \right|$.
+We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
 
@@
 \begin{equation}
-  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecb, \vecvv)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
   \label{eq:retro_microfacet_model}.
@@
 In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
+Note that for this we impose one technical assumption on $D$:
+we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
 
@@
 \end{equation}
-As anisotropic Phong, Beckmann, and GGX distributions are reflection-symmetric around the normal, i.e.
-they take the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using them with the back-vector obeys symmetry.
+As $D$ is reflection-symmetric around the normal, i.e.
+it takes the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with the back-vector obeys symmetry.
 Further, the reflection symmetry of $D$ along with equation \eqref{eq:nk} gives
@@
   \vdot{\vecll}{\vech(\vecll,\vecv)}.
+\end{equation}\label{eq:F}
+As the Fresnel term takes the form $F(\vecx, \vecy) = F\left(\vdot{\vecx}{\vecy}\right)$, we can conclude
+  \begin{equation}
+    F\left((\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =
+    F\left(\vdot{\vecll}{\vech(\vecll,\vecv)}\right) = F\left(\vecll, \vech(\vecll,\vecv)\right).
 \end{equation}    
-Using this along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
+Further, using equation \eqref{eq:kh} along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
 \begin{equation}
@@
   f(\vecvv, \vecl) &=
-  \frac{D(\vech(\vecvv, \vecl)) G_{2}(G_1(\vecvv, \vecl), G_1(\vecl, \vecvv))}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
+  \frac{D(\vech(\vecvv, \vecl)) G_{2}(G_1(\vecvv, \vecl), G_1(\vecl, \vecvv)) F(\vecvv, \vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
   &=
-  \frac{D(\vech(\vecll, \vecv)) G_{2}(G_1(\vecv, \vecll), G_1(\vecll, \vecv))}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} } =
+  \frac{D(\vech(\vecll, \vecv)) G_{2}(G_1(\vecv, \vecll), G_1(\vecll, \vecv)) F(\vecll, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} }\\
+  &=
   f(\vecll, \vecv) =
@@
 \item Pseudo-code is provided in Listing~\ref{lst:implementation}.
+  \todo{mraab: I think source code may be overkill, given the simplicity of the model.}
 \end{itemize}

### 2025-07-15 article.tex
WIP

--- 2025-07-02 article.tex
@@
 For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
-We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets are these requirements.
+We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements.
 It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
@@
 i.e. we get the albedo of the regular microfacet BRDF for the reflected direction.
-As the original microfacet BRDF fulfills energy conversion, the retro-reflective modification does, too.
-%% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
-%% \begin{equation}
-%%   \begin{split}
-%%   \rho(\vecvv) &=
-%%   \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
-%%   &\leq
-%%   \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
-%%   = 1.
-%%   \end{split}
-%% \end{equation}
+Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
+\begin{equation}
+  \begin{split}
+  \rho(\vecvv) &=
+  \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
+  &\leq
+  \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
+  = 1.
+  \end{split}
+\end{equation}
 
@@
 Matthias Raab \newline
-\href{mailto:mraab@nvidia.com}{mraab@nvidia.com} \newline
+\href{mailto:matthias.raab@gmail.com}{matthias.raab@gmail.com} \newline
 &

### 2025-07-15 scripts/retro.py

@@ -20,11 +20,11 @@ def D_ggx(M, r):
     Ddenom = math.pi * ax * ay * sqr(sqr(M[0]/ax) + sqr(M[1]/ay) + sqr(M[2]))
     return 1.0 / max(Ddenom, DENOM_TOLERANCE)
 
-
 def ggx_lambda(V, r):
     if abs(V[2]) < FLT_EPSILON:
         return 0.0
-    return (-1.0 + sqrt(1.0 + (sqr(r*V[0]) + sqr(r*V[1]))/sqr(V[2]))) / 2.0
+    a = sqr(r)
+    return (-1.0 + sqrt(1.0 + (sqr(a*V[0]) + sqr(a*V[1]))/sqr(V[2]))) / 2.0
 
 def G2_ggx(V, L, m, r):
     return 1.0 / (1.0 + ggx_lambda(V, r) + ggx_lambda(L, r))

### 2025-07-22 article.tex
(mostly) extend for refraction on microfacets (BTDF)

--- 2025-07-10 article.tex
@@
 \newcommand{\vech}{\mathbf{h}}
+\newcommand{\vechr}{\mathbf{h_r}}
+\newcommand{\vecht}{\mathbf{h_t}}
 \newcommand{\vecb}{\mathbf{b}}
+\newcommand{\vecbr}{\mathbf{b_r}}
+\newcommand{\vecbt}{\mathbf{b_t}}
 \newcommand{\vecx}{\mathbf{x}}
+\newcommand{\vecxx}{\mathbf{x'}}
 \newcommand{\vecy}{\mathbf{y}}
+\newcommand{\vecyy}{\mathbf{y'}}
+\newcommand{\etav}{\eta_{\mathbf{v}}}
+\newcommand{\etal}{\eta_{\mathbf{l}}}
+\newcommand{\etax}{\eta_{\mathbf{x}}}
+\newcommand{\etay}{\eta_{\mathbf{y}}}
 \newcommand{\vdot}[2]{\left\langle #1, #2 \right\rangle}
+%\newcommand{\vdot}[2]{#1\cdot#2}
+\newcommand{\avdot}[2]{\lvert\vdot{#1}{#2}\rvert}
 
@@
   \small
-  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BRDF implementation to produce a visually plausible, energy-conserving retroreflective result.
+  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
   \end{abstract}
@@
 
-In this document, we detail the correct usage and derivation of this retro-reflective BRDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
-We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BRDF is fully plausible (Section~\ref{sec:plausibility}).
+In this document, we detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
+We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}).
 Finally, we sketch how the model can be implemented in practice (Section~\ref{sec:implementation}).
 
-\section{Microfacet BRDF Models}
+\section{Microfacet BSDF Models}
 \label{sec:microfacets}
@@
       \hline
-      $\vech$ & half vector, $\vech = \vech(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
+      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
+      \hline
+      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to \\
       \hline
@@
       \hline
-      $\Omega$ & hemisphere of directions around the normal \\
+      $\Omega$ & hemisphere of directions around the normal \todo{need more formalization for transmission}\\
       \hline
@@
 
-Using the notations from figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
+Using the notations from Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
 \begin{equation}
-  f(\vecv, \vecl) = \frac{D(\vech) \, G_{2}(\vecv,\vecl) \, F(\vech, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
-  \label{eq:microfacet_model}
+  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
+  \label{eq:microfacet_brdf}
 \end{equation}
-where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
-The half-vector $\vech$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacet with normals aligned with the half-vector will reflect light).
-In Section~4.1 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equation~\ref{eq:microfacet_model}.
-
+where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vechr) \vdot{\vecn}{\vechr} d\vechr = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
+The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
+Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
+\begin{equation}
+  f_t(\vecv, \vecl) =
+  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
+       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecv}{\vecht} + \etal \avdot{\vecl}{\vecht} \right)^2}
+  \label{eq:microfacet_btdf}.
+\end{equation}
+In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
 In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector.
@@
 \end{equation}
-Mathematically, the BRDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
+Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
 $G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
@@
 \begin{equation}
-  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\vdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
+  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
 \end{equation}
@@
   G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
-  \min\left\lbrace\frac{2\vdot{\vecv}{\vecn}\vdot{\vecn}{\vech}}{\vdot{\vecv}{\vech}}, 1\right\rbrace,
+  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
 \end{equation}
@@
 
-Instead of half-vector $\vech$, we can use the back-vector $\vecb$~\cite{BelcourBackvector} defined as
+Instead of reflection half-vector $\vechr$, we can use the reflection back-vector $\vecbr$~\cite{BelcourBackvector} defined as
 \begin{equation}
-  \vecb(\vecv, \vecl) = \vech(\vecvv, \vecl) = \frac{\vecvv + \vecl}{\lVert \vecvv + \vecl \rVert}, \text{ with } \vecvv = \text{reflect}(\vecv, \vecn),
+  \vecbr(\vecv, \vecl) = \vechr(\vecvv, \vecl) = \frac{\vecvv + \vecl}{\lVert \vecvv + \vecl \rVert}, \text{ with } \vecvv = \text{reflect}(\vecv, \vecn),
 \end{equation}
 to select the reflecting microfacet. %, we are to re-derive the microfacet model.
+For refraction, we can construct a similar refraction back-vector $\vecbt(\vecv, \vecl) = \vecht(\vecvv, \vecl)$.
 
@@
     \includegraphics[width=0.4\textwidth]{figures/backvector.pdf}
+    \includegraphics[width=0.4\textwidth]{figures/backvector_refract.pdf}
   \end{center}
-  \caption{Geometry of the back vector, illustrated in two dimensions.}
+  \caption{Geometry of the back-vector for reflection and refraction, illustrated in two dimensions.}
   \label{fig:backvector}
@@
 \end{equation}
+\todo{Formulate this for refraction?}
 
 \paragraph*{Microfacet reflection term.}
-There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements.
+There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
 As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead.
@@
 We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
+\todo{Formulate this for refraction?}
 
@@
 We simply apply Smith or v-cavities masking to the mirrored microfacet distribution.
-With Smith masking, this is actually identical to a regular microfacet BRDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
+With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
 For v-cavities masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
@@
 \begin{equation}
-  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  f_\text{retro}(\vecv, \vecl) = f_{r, \text{retro}}(\vecv, \vecl) + f_{t, \text{retro}}(\vecv, \vecl) =
+  f_r(\vecvv, \vecl) + f_t(\vecvv, \vecl)
   \label{eq:retro_microfacet_model}.
 \end{equation}
+\todo{separte BRDF and BTDF}
 In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
@@
 \end{equation}
-For the back-vector we show $\vech(\vecvv, \vecl) = \text{reflect}\left(\vech(\vecll, \vecv), \vecn\right)$, as
+For the reflection back-vector we show $\vechr(\vecvv, \vecl) = \text{reflect}\left(\vechr(\vecll, \vecv), \vecn\right)$, as
 \begin{equation}
@@
 and therefore $\lVert \vecvv + \vecl \rVert = \lVert \vecll + \vecv \rVert$ as well.
+Similarly, for refraction we have
+\begin{equation}
+  \begin{split}
+  \etav\vecvv + \etal\vecl &=
+  \etav\left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \etal\vecl \\
+  &= \etav\left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \etal\left(- \vecll + 2 \vdot{\vecll}{\vecn} \vecn \right)\\
+  &= -(\etal\vecll + \etav\vecv) + 2 \vdot{(\etal\vecll + \etav\vecv)}{\vecn} \vecn =
+  \text{reflect}(\etal\vecll + \etav\vecv, \vecn)
+  \end{split}
+\end{equation}
+which implies $\lVert \etav\vecvv + \etal\vecl \rVert = \lVert \etal\vecll + \etal\vecv \rVert$ and
+results in $\vecht(\vecvv, \vecl) = \text{reflect}\left(\vecht(\vecll, \vecv), \vecn\right)$.
 From that we can conclude
@@
     \vdot{\mathbf{t_2}}{\vech(\vecvv, \vecl)} &=
-    -\vdot{\mathbf{t_2}}{\vech(\vecll, \vecv)}.\\
+    -\vdot{\mathbf{t_2}}{\vech(\vecll, \vecv)}\\
     \end{split}
 \end{equation}
+for both reflection ($\vech = \vechr$) and refraction ($\vech = \vecht$) cases.
 As $D$ is reflection-symmetric around the normal, i.e.
-it takes the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with the back-vector obeys symmetry.
+it takes the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
 Further, the reflection symmetry of $D$ along with equation \eqref{eq:nk} gives
@@
 \begin{equation}
-  \vdot{\vecvv}{\vecl} = -\vdot{\vecv}{\vecl} + 2 \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn}
-  = \vdot{\vecv}{-\vecl + 2 \vdot{\vecl}{\vecn}\vecn} = \vdot{\vecv}{\vecll}
+  \begin{split}
+%  \vdot{\vecx}{\vecy} = -\vdot{\vecxx}{\vecy} + 2 \vdot{\vecxx}{\vecn}\vdot{\vecy}{\vecn}
+%  = \vdot{\vecxx}{-\vecyy + 2 \vdot{\vecyy}{\vecn}\vecn} = \vdot{\vecxx}{\vecyy}
+    \vdot{\vecxx}{\vecyy} &= \vdot{-\vecx}{\vecyy} + 2 \vdot{\vecx}{\vecn}\vdot{\vecn}{\vecyy}\\
+  &= \vdot{-\vecx}{-\vecy} + \vdot{-\vecx}{2 \vdot{\vecy}{\vecn}\vecn} + 2 \vdot{\vecx}{\vecn}\vdot{\vecn}{\vecy}\\
+  &= \vdot{\vecx}{\vecy}
+\end{split}
 \end{equation}
 and therefore
-\begin{equation}\label{eq:kh}
-  \vdot{\vecvv}{\vech(\vecvv, \vecl)} =
-  \frac{1 + \vdot{\vecl}{\vecvv}}{\lVert \vecvv + \vecl \rVert} =
-  \frac{1 + \vdot{\vecll}{\vecv}}{\lVert \vecll + \vecv \rVert} =
-  \vdot{\vecll}{\vech(\vecll,\vecv)}.
-\end{equation}\label{eq:F}
+\begin{equation}\label{eq:khr}
+  \vdot{\vecx}{\vech(\vecx, \vecy)} =
+  \frac{1 + \vdot{\vecy}{\vecx}}{\lVert \vecx + \vecy \rVert} =
+  \frac{1 + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =
+  \vdot{\vecyy}{\vech(\vecyy,\vecxx)}
+\end{equation}
+as well as
+\begin{equation}\label{eq:kht}
+  \vdot{\vecx}{\vecht(\vecx, \vecy)} =
+  \frac{-\etax  -\etay \vdot{\vecx}{\vecy}}{\lVert -\etax\vecx  -\etay\vecy \rVert} =
+  \frac{-\etax \vdot{\vecxx}{\vecxx} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
+  \vdot{\vecxx}{\vech(\vecyy,\vecxx)}.
+\end{equation}
 As the Fresnel term takes the form $F(\vecx, \vecy) = F\left(\vdot{\vecx}{\vecy}\right)$, we can conclude
@@
 \end{equation}    
-Further, using equation \eqref{eq:kh} along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
+Further, using equations \eqref{eq:khr} and \eqref{eq:kht} along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
 \begin{equation}
@@
 and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the v-cavity masking function.
-In summary, we can conclude the symmetry of the BRDF
+In summary, we can conclude the symmetry of both the BRDF
 \begin{equation}
   \begin{split}
-  f_{\text{retro}}(\vecv, \vecl) =
-  f(\vecvv, \vecl) &=
+  f_{r,\text{retro}}(\vecv, \vecl) =
+  f_r(\vecvv, \vecl) &=
   \frac{D(\vech(\vecvv, \vecl)) G_{2}(G_1(\vecvv, \vecl), G_1(\vecl, \vecvv)) F(\vecvv, \vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
@@
   &=
-  f(\vecll, \vecv) =
-  f_{\text{retro}}(\vecl, \vecv).
+  f_r(\vecll, \vecv) =
+  f_{r,\text{retro}}(\vecl, \vecv)
+  \end{split}
+\end{equation}
+and the BTDF
+\begin{equation}
+  \begin{split}
+    &f_{t,\text{retro}}(\vecv, \vecl)
+    =  f_t(\vecvv, \vecl)\\
+    &= \frac{\avdot{\vecvv}{\vecht(\vecvv, \vecl)}\avdot{\vecl}{\vecht(\vecvv, \vecl)}\etav^2 D(\vecht(\vecvv, \vecl)) \, G_{2}(\vecvv,\vecl) \, \left(1 - F(\vecvv, \vecht(\vecvv, \vecl)) \right)}
+    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\
+    &= \frac{\avdot{\vecv}{\vecht(\vecv, \vecll)}\avdot{\vecll}{\vecht(\vecv, \vecll)}\etav^2 D(\vecht(\vecll, \vecv)) \, G_{2}(\vecll,\vecv) \, \left(1 - F(\vecll, \vecht(\vecll, \vecv)) \right)}
+    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\
+    &=
+    \frac{\etav^2}{\etal^2}f_t(\vecll, \vecv) =
+    \frac{\etav^2}{\etal^2}f_{t,\text{retro}}(\vecl, \vecv).
   \end{split}
@@
 
-Given an implementation of a regular microfacet BRDF, extending it to retro-reflection is straightforward:
+Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is straightforward:
 
@@
 \item
-  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BRDF.
+  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
   This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
 \item
-  As the albedos of standard BRDF and retro-reflective BRDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
+  As the albedos of standard BSDF and retro-reflective BSDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
 \item Pseudo-code is provided in Listing~\ref{lst:implementation}.

### 2025-07-23 article.tex
further cleanup for refraction, fix some formulas and notation

@@
       \hline
-      $\vecn$ & surface normal \\
+      $\vecn$ & surface normal, with the convention that it points to
+      the medium with lower index of refraction in cases dealing with $\vecv$ and $\vecl$ pointing to different sides\\
       \hline
@@
       \hline
-      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to \\
+      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to (always points to medium with lower index of refraction)\\
       \hline
@@
 $\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
+%      \hline
+%      $\Omega$ & hemisphere of directions around the normal\\
       \hline
-      $\Omega$ & hemisphere of directions around the normal \todo{need more formalization for transmission}\\
+      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
       \hline
@@
 \end{equation}
-where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vechr) \vdot{\vecn}{\vechr} d\vechr = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
+where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
 The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
@@
   \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
-       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecv}{\vecht} + \etal \avdot{\vecl}{\vecht} \right)^2}
+       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
   \label{eq:microfacet_btdf}.
@@
 \begin{equation}\label{eq:visible}
-  \vdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
+  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
 \end{equation}
@@
 \begin{equation}
-  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
+  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
 \end{equation}
@@
 
-As illustrated in figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
+As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
 Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
@@
 \end{equation}
-\todo{Formulate this for refraction?}
+\todo{Formulate this for refraction or remove/simplify?}
 
@@
 We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
-\todo{Formulate this for refraction?}
+\todo{Formulate this for refraction or remove/simplify?}
 
@@
 The final retro-reflective microfacet model can be written as
-\begin{equation}
+\begin{equation} \label{eq:retro_microfacet_model}
   %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
-  f_\text{retro}(\vecv, \vecl) = f_{r, \text{retro}}(\vecv, \vecl) + f_{t, \text{retro}}(\vecv, \vecl) =
-  f_r(\vecvv, \vecl) + f_t(\vecvv, \vecl)
-  \label{eq:retro_microfacet_model}.
+  f_\text{retro}(\vecv, \vecl) =
+  \begin{cases}
+    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if} \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn} \geq 0, \\
+    f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
+  \end{cases}
 \end{equation}
-\todo{separte BRDF and BTDF}
 In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
@@
 
-In analogy to $\vecvv$ we let $\vecll := \text{reflect}(\vecl, \vecn)$.
+In analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
 First we observe
@@
 As $D$ is reflection-symmetric around the normal, i.e.
-it takes the form $D(\vech) = D\left(\vdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
-Further, the reflection symmetry of $D$ along with equation \eqref{eq:nk} gives
+it takes the form $D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
+Further, the reflection symmetry of $D$ along with Equation \eqref{eq:nk} gives
 \begin{equation}
@@
   G_{1,\text{Smith}}(\vecv) &=
-  \frac{\vdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
-  &= \frac{\vdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =
+  \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
+  &= \frac{\avdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecvv}: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =
   G_{1,\text{Smith}}(\vecvv)
@@
   \frac{1 + \vdot{\vecy}{\vecx}}{\lVert \vecx + \vecy \rVert} =
-  \frac{1 + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =
+  \frac{\vdot{\vecyy}{\vecyy} + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =
   \vdot{\vecyy}{\vech(\vecyy,\vecxx)}
@@
   \frac{-\etax  -\etay \vdot{\vecx}{\vecy}}{\lVert -\etax\vecx  -\etay\vecy \rVert} =
-  \frac{-\etax \vdot{\vecxx}{\vecxx} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
-  \vdot{\vecxx}{\vech(\vecyy,\vecxx)}.
+  \frac{-\etax \vdot{\vecyy}{\vecyy} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
+  \vdot{\vecyy}{\vech(\vecyy,\vecxx)}.
 \end{equation}
@@
 \end{equation}    
-Further, using equations \eqref{eq:khr} and \eqref{eq:kht} along with equations \eqref{eq:nh} and \eqref{eq:nk} we get
+Further, using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get
 \begin{equation}
@@
 and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the v-cavity masking function.
-In summary, we can conclude the symmetry of both the BRDF
+In summary, we can conclude the symmetry of the BRDF
 \begin{equation}
@@
   f_r(\vecll, \vecv) =
-  f_{r,\text{retro}}(\vecl, \vecv)
+  f_{r,\text{retro}}(\vecl, \vecv).
   \end{split}
 \end{equation}
-and the BTDF
+Analogously, we obtain the symmetry requirement of the BTDF
 \begin{equation}
@@
     &= \frac{\avdot{\vecvv}{\vecht(\vecvv, \vecl)}\avdot{\vecl}{\vecht(\vecvv, \vecl)}\etav^2 D(\vecht(\vecvv, \vecl)) \, G_{2}(\vecvv,\vecl) \, \left(1 - F(\vecvv, \vecht(\vecvv, \vecl)) \right)}
-    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\
+    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecvv}{\vecht(\vecvv, \vecl)} + \etal \vdot{\vecl}{\vecht(\vecvv, \vecl)}\right)^2} \\
     &= \frac{\avdot{\vecv}{\vecht(\vecv, \vecll)}\avdot{\vecll}{\vecht(\vecv, \vecll)}\etav^2 D(\vecht(\vecll, \vecv)) \, G_{2}(\vecll,\vecv) \, \left(1 - F(\vecll, \vecht(\vecll, \vecv)) \right)}
-    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \avdot{\vecvv}{\vecht} + \etal \avdot{\vecl}{\vecht}\right)^2} \\
+    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \vdot{\vecv}{\vecht(\vecv, \vecll)} + \etal \vdot{\vecll}{\vecht(\vecv, \vecll)}\right)^2} \\
     &=
@@
 \begin{equation}
-  \rho_{\text{retro}}(\vecv) =
-  \int_\Omega f_{\text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
-  \int_\Omega f(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho(\vecvv)
+  \begin{split}
+  \rho_{r, \text{retro}}(\vecv) =
+  \int_{\Omega_{\vecv}} f_{r, \text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
+  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho_r(\vecvv),\\
+  \rho_{t, \text{retro}}(\vecv) =
+  \int_{\Omega_{-\vecv}} f_{t, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} d\vecl =
+  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} d\vecl = \rho_t(\vecvv),
+  \end{split}
 \end{equation}
-i.e. we get the albedo of the regular microfacet BRDF for the reflected direction.
-As the original microfacet BRDF fulfills energy conversion, the retro-reflective modification does, too.
+i.e. we get the albedo of the regular microfacet BSDF for the reflected direction.
+As the original microfacet BSDF fulfills energy conversion, the retro-reflective modification does, too.
 %% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
@@
 \item
-  As the albedos of standard BSDF and retro-reflective BSDF are essentially identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
+  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
 \item Pseudo-code is provided in Listing~\ref{lst:implementation}.
-  \todo{mraab: I think source code may be overkill, given the simplicity of the model.}
 \end{itemize}
@@
     language=GLSL,
-    caption=Pseudo code to implement the retro-reflective microfacet model.,
+    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
     label=lst:implementation,
@@
     ]
-vec3 brdf_retro(vec3 L, vec3 V, vec3 N, vec3 X, vec3 Y)
+Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
 {
-    vec3 Vp = -reflect(V,N); // jp - is this wrong? Why the minus sign
+    vec3 Vp = -V + 2.0 * dot(V, N) * N;
+    return sample_bsdf(V, L, N, params);
+}
+Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
+{
+    vec3 Vp = -V + 2.0 * dot(V, N) * N;
+    return eval_bsdf(V, L, N, params);
+}
 
-    // The following could all be replaced with:
-    // We could give the simplified form of brdf_ggx, but not essential.
-    return brdf_ggx(L, Vp, N, X, Y);
-
-    /*
-    float NdotL = dot(N, L);
-    float NdotV = dot(N, Vp);
-    if (NdotL < 0 || NdotV < 0) return vec3(0);
-
-    vec3 B = normalize(L+Vp);
-    float NdotH = dot(N, B);
-    vec3 wB = vec3(dot(B,X),  dot(B,Y),  dot(B,N));
-    vec3 wL = vec3(dot(L,X),  dot(L,Y),  dot(L,N));
-    vec3 wV = vec3(dot(Vp,X), dot(Vp,Y), dot(Vp,N));
-
-    float D = D_GGX(wB, ax, ay);
-    float G = G2_GGX(wV, wL, ax, ay);
-
-    float c = dot(L, B);
-    float F = schlick(f0, c); // I don't think we need to give the form of the Fresnel
-
-    float val = Ks * D * G * F / (4 * NdotL * NdotV);
-    return vec3(val);
-    */
-}
 \end{lstlisting}

### 2025-08-28 article.tex
Adding refractive equations
Updating article with refractive equations in Section 3.

@@
 \paragraph*{Jacobian of Back-Vector.}
-First we observe that  Jacobian of the change of variable from the back-vector to the light vector $\vecl$~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
+First we observe that Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
 \begin{equation}
-\left\lVert \dfrac{\partial \omega_{\vecb}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
+\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
 \end{equation}
-\todo{Formulate this for refraction or remove/simplify?}
+We also observe that this is true for refractive microfacets~(Equation (17) of \citet{WalterMicrofacet}):
+\begin{equation}
+\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2}
+\end{equation}
+This is due to the fact that we have to differentiate with respect to the light direction (which is not altered by the back vector).
 
@@
 \begin{equation}
-  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb, \vecm) }{4 \left| \vdot{\vecvv}{\vecb} \right|^2}.
+  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
 \end{equation}
 We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
-\todo{Formulate this for refraction or remove/simplify?}
+Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
+\begin{equation}
+  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
+\end{equation}
 

### 2025-09-12 article.tex
Adding previous work section

@@
 Rendering of retro-reflective materials is useful in a variety of contexts.
-A number of CG models have been proposed.
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
-We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets are these requirements.
-It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
+For security garnments and signs, retroreflective appearances are heavily used.
+\todo{Different usage and forms} We known for measurements that retro-reflective tapes exhibit an imperfect (rough) retro-reflective lobe~\cite{BelcourBackvector}.
 
-In this document, we detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
+\paragraph{Previous works.}
+Modeling of retro-reflectivity in the surface scattering function has received little attention so far.
+For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets.
+However, their model only provide pure retro-reflection which would lack artistic control.
+They later extented this model account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
+However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
+
+\paragraph{Our model.} In this document, we present a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets production requirements.
+It is also a simple modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
+
+We detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
 We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}).
@@
   \end{center}
-  \caption{Notations used.}
+  \caption{
+    Notations used.
+  }
   \label{fig:notation}

### 2025-09-26 article.tex
note that albedos are identical and that BTDF/BRDF combinations are possible for forwared and retro reflective variants

@@
 %% \end{equation}
+We further note that under our requirement of a reflection-symmetric microfacet normal distribution, the geometric
+configurations for $\vecv$ and $\vecvv$ are identical.
+Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retro-reflection are the same, i.e.
+$\rho_{r, \text{retro}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{retro}}(\vecv) = \rho_t(\vecv)$.
 
@@
 
+\paragraph*{Combination of BRDF and BTDF}
+
+The albedo is identical for forward- and retro-reflective BRDF as well as BTDF,
+which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
+When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+
 \bibliographystyle{jcgt}

### 2025-10-30 article.tex

@@
 Rendering of retro-reflective materials is useful in a variety of contexts.
-For security garnments and signs, retroreflective appearances are heavily used.
-\todo{Different usage and forms} We known for measurements that retro-reflective tapes exhibit an imperfect (rough) retro-reflective lobe~\cite{BelcourBackvector}.
+For security garments and signs, retroreflective appearances are heavily used.
+A number of CG models have been proposed.
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
+We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements.
+It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
 

### 2025-12-19 article.tex
Add updated figures, and improvements to text.

@@
 \setciteauthor{Matthias Raab and Laurent Belcour and Frankie Liu and Jamie Portsmouth}
-\setcitetitle{The Minimal Retroreflective Microfacet Model (MRMM)}
+\setcitetitle{The Minimal Retroreflective Microfacet Model (MRM)}
 %\setheadtitle{Abbreviated title, only if full title won't fit in page headers}
@@
 \captionsetup[table]{skip=10pt}
+
+\lstdefinestyle{snippet}{
+  float=tp,
+  floatplacement=tbp,
+  abovecaptionskip=5pt
+}
 
@@
   \vspace*{-0.5cm}
-  \centering
-    \includegraphics[width=0.32\linewidth]{figures/sphere_side_r0.3.png     }
-    \includegraphics[width=0.32\linewidth]{figures/sphere_quarter_r0.3.png  }
-    \includegraphics[width=0.32\linewidth]{figures/sphere_front_r0.3.png    }
-  \caption{Comparison of our retroreflector model (top) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (bottom).
-The lighting varies (left to right) from side-on, to quarter-on, and straight-on, exhibiting the strong retro-reflective peak.
+    \hfill
+    \includegraphics[width=0.25\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
+    \includegraphics[width=0.25\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
+    \includegraphics[width=0.25\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
+    \vspace{0.06cm}
+    \includegraphics[width=0.25\linewidth]{figures/cone/retro_cone_cropped_label.png}
+    \includegraphics[width=0.25\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
+    \includegraphics[width=0.25\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
+    \hfill
+  \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
+The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak.}
    \vspace{-0.5cm}
-  \label{fig:teaser}}
+  \label{fig:teaser}
 }
@@
   \small
-  We introduce the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
+  We describe the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
   \end{abstract}
@@
 Rendering of retro-reflective materials is useful in a variety of contexts.
-For security garments and signs, retroreflective appearances are heavily used.
-A number of CG models have been proposed.
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood microfacet models.
+For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some examples).
+Such materials are typically designed to be retroreflective via a substructure of elements which preferentially scatter light backwards.
+A number of CG models have been proposed to model retro-reflective materials, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.
+
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
 We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements.
@@
 
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
 \paragraph{Previous works.}
@@
 For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets.
-However, their model only provide pure retro-reflection which would lack artistic control.
-They later extented this model account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
+However, their model only provides pure retro-reflection which would lack artistic control.
+They later extented this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
 However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
@@
 We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}).
-Finally, we sketch how the model can be implemented in practice (Section~\ref{sec:implementation}).
+In Section~\ref{sec:implementation}, we sketch how the model can be implemented in practice, which is done trivially by reusing existing microfacet BSDF implementations.
+Finally, we provide some visual results and discussion of the behavior (Section~\ref{sec:results}).
 
@@
   \caption{
-    Notations used.
+    Notation used.
   }
@@
 
-Using the notations from Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
+Using the notation of Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
 \begin{equation}
@@
 \begin{equation}
-G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right)
+G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right) \ .
 \end{equation}
@@
 \begin{equation}\label{eq:visible}
-  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
+  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \,G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
 \end{equation}
@@
 \end{equation}
-where analytic expressions are available for Beckmann and GGX distributions, and the v-cavity masking function
+where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function
 \begin{equation}
@@
 
-As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as an reflection of the view direction on a plane aligned with $\vecv$ and perpendicular to the surface.
-Intuitively, this is equivalent of observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
-While there is no immediate physical foundation to this construction, this interpretation can serve as motivation why components from the regular microfacet model still work when changing the domain to the back-vector.
+As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual'' plane aligned with $\vecv$ and perpendicular to the surface.
+Intuitively, this is equivalent to observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
+While there is no immediate physical foundation to this construction, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
+We now detail the mathematical consequences of using the back-vector as microfacet selector.
 
 \paragraph*{Jacobian of Back-Vector.}
-First we observe that Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
+First we observe that the Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
 \begin{equation}
-\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|}
+\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|} \ .
 \end{equation}
@@
 
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=0.99\linewidth]{figures/brdf_plots.pdf}
+  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
+The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$.
+The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
+  \label{fig:brdf_plots}
+\end{figure}
+
 \paragraph*{Microfacet reflection term.}
@@
 As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead.
-It follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
+From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
 \begin{equation}
@@
 \paragraph*{Final form.}
-The final retro-reflective microfacet model can be written as
+It follows from all this that the final retro-reflective microfacet model can be written simply as
 \begin{equation} \label{eq:retro_microfacet_model}
@@
 we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
+
+Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue).
+For both MRM and GGX, the Fresnel factor is set to 1.
+The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
+It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
+The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane.
+In this case.
+both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retro-reflection direction respectively for GGX and MRM).
 
@@
   &= -(\vecll + \vecv) + 2 \vdot{(\vecll + \vecv)}{\vecn} \vecn =
-  \text{reflect}(\vecll + \vecv, \vecn)
+  \text{reflect}(\vecll + \vecv, \vecn) \ .
   \end{split}
@@
   &= -(\etal\vecll + \etav\vecv) + 2 \vdot{(\etal\vecll + \etav\vecv)}{\vecn} \vecn =
-  \text{reflect}(\etal\vecll + \etav\vecv, \vecn)
+  \text{reflect}(\etal\vecll + \etav\vecv, \vecn) \ .
   \end{split}
@@
     \vdot{\vecn}{\vech(\vecvv, \vecl)} &=
-    \vdot{\vecn}{\vech(\vecll, \vecv)}\\
+    \vdot{\vecn}{\vech(\vecll, \vecv)}  \ , \\
     \vdot{\mathbf{t_1}}{\vech(\vecvv, \vecl)} &=
-    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)}\\
+    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)} \ , \\
     \vdot{\mathbf{t_2}}{\vech(\vecvv, \vecl)} &=
@@
 for both reflection ($\vech = \vechr$) and refraction ($\vech = \vecht$) cases.
-As $D$ is reflection-symmetric around the normal, i.e.
-it takes the form $D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)$, we can thus infer that using it with a back-vector obeys symmetry.
+Since we assumed $D$ is reflection-symmetric around the normal, i.e. it takes the form
+\begin{equation}
+D(\vech) = D\left(\avdot{\vecn}{\vech}, \lvert \vdot{\mathbf{t_1}}{\vech} \rvert, \lvert \vdot{\mathbf{t_2}}{\vech} \rvert\right)
+\end{equation}
+we can thus infer that using it with a back-vector obeys symmetry.
 Further, the reflection symmetry of $D$ along with Equation \eqref{eq:nk} gives
@@
   \begin{equation}
-    F\left((\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =
+    F\left(\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =
     F\left(\vdot{\vecll}{\vech(\vecll,\vecv)}\right) = F\left(\vecll, \vech(\vecll,\vecv)\right).
@@
 \end{equation}
-and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the v-cavity masking function.
-In summary, we can conclude the symmetry of the BRDF
+and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the V-cavity masking function.
+In summary, we can conclude the reciprocity symmetry of the BRDF
 \begin{equation}
@@
 \end{equation}
-Analogously, we obtain the symmetry requirement of the BTDF
+Analogously, we obtain the reciprocity symmetry requirement of the BTDF
 \begin{equation}
@@
 \end{equation}
-For the albedo we simply have
+For the directional albedos we simply have
 \begin{equation}
@@
 
-Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is straightforward:
+Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward:
 
@@
 
-\noindent\begin{minipage}[tb]{\linewidth}
+\noindent
 \begin{lstlisting}[
+    style=snippet,
     language=GLSL,
@@
 }
+\end{lstlisting}
 
-\end{lstlisting}
-\end{minipage}
+\section{Results and Discussion}
+
+\label{sec:results}
+
+\paragraph*{Incorporation in a standard microfacet model}
+
+Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retro-reflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.
+
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_0_cropped.png}
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_45_cropped.png}
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_90_cropped.png} \\
+    \vspace{0.06cm}
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_0_cropped.png}
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_45_cropped.png}
+    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
+  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction.
+The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).}
+  \label{fig:conductor_renders}
+\end{figure}
 
@@
 
-The albedo is identical for forward- and retro-reflective BRDF as well as BTDF,
-which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
+The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
 When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+
+Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
+Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
+    \vspace{0.06cm}
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
+    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
+  \caption{Renders of a glass shaderball illuminated from the side.
+The index of refraction is $1.0$ (top) or $1.5$ (bottom).
+From left to right: classic microfacet BSDF (left), retro-reflective BSDF (middle), and combined retro-reflective BRDF and classic BTDF (right).
+}
+  \label{fig:dielectric_renders}
+\end{figure}
+
+\section{Conclusion}
+
+We have presented a retro-reflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
+The model is straightforward to implement in existing renderers supporting microfacet BSDFs, as it only requires reflecting the view direction about the surface normal before evaluating the standard microfacet model.
+We have shown that the model fulfills symmetry and energy conservation requirements, making it a plausible microfacet model.
+The retro-reflective microfacet model can be used to represent materials with strong retro-reflective properties, such as certain fabrics or retro-reflective coatings.
+
+\paragraph{Further work.}
+For future work it would be interesting to further validate the MRM retro-reflective microfacet model against measured retro-reflective materials.
 

### 2025-12-19 code/brdf_calculator.cpp
Add code for BRDF evaluation/comparison plot.

@@ -1,0 +1,103 @@
+/*
+Prerequisites:
+
+- Clone the glm repository from https://github.com/g-truc/glm
+- Set the path to the glm repository in the include directive below
+
+General compilation and running commands:
+
+g++ -std=c++17 -O3 -o brdf_calculator brdf_calculator.cpp -I/path/to/glm
+./brdf_calculator
+
+Compilation and running commands for Peter's setup:
+
+cd ~/repos/EON_jcgt_paper/code
+g++ -std=c++17 -O3 -o brdf_calculator brdf_calculator.cpp -I/Users/pkutz/repos/glm/
+./brdf_calculator
+
+Compilation and running script for Peter's setup (modify for your own setup):
+
+./plot_brdfs.sh
+*/
+
+#include "brdf_common.h"
+
+#include <fstream>
+#include <string>
+#include <vector>
+
+class BRDFCalculator {
+public:
+    void computeAndOutputBRDFData() {
+        vec3 rho = vec3(1.f);  // Albedo.
+
+        std::vector<float> roughnesses = {0.1f, 0.666f}; // Roughness values.
+
+        for (float roughness : roughnesses)
+        {
+            std::vector<float> thetas_i = {45.0f, 80.0f}; // Incident zenith angles in degrees.
+            std::vector<float> phis = {0.0f, 45.0f};     // Relative azimuth angles in degrees.
+
+            // Loop through each combination of theta_i and phi to generate separate CSV files.
+            for (float theta_i : thetas_i) {
+                for (float phi_o : phis) {
+                    std::string filename = "brdf_theta_i_" + std::to_string((int)theta_i) + "_phi_" + std::to_string((int)phi_o) + "_roughness_" + std::to_string((roughness)) + ".csv";
+                    std::ofstream file(filename);
+
+                    // Write headers to the CSV file.
+                    file << "theta_o,Lambert,ON,QON,footnoteQON,fujiiQON,FON,EON,GGX,MRM\n";
+
+                    float angle_tol = 1.0e-4f;
+                    // Compute BRDF values over the range of theta_o.
+                    for (float theta_o = -90.0f+angle_tol; theta_o <= 90.0f-angle_tol; theta_o += 1.0f) {
+                        vec3 wi_local = calculate_wi_local(theta_i);
+                        vec3 wo_local = calculate_wo_local(theta_o, phi_o);
+
+                        // Call each BRDF function with the generated directions.
[…]

### 2025-12-19 code/brdf_common.h

@@ -1,0 +1,88 @@
+#include <chrono>
+#include <cmath>
+#include <iostream>
+#include <random>
+
+#define GLM_FORCE_CTOR_INIT
+#include <glm/glm.hpp>
+
+using namespace glm; // This allows the GLSL code to be used directly
+
+#include "brdf_implementations.glsl"
+
+class Timer {
+private:
+    std::chrono::high_resolution_clock::time_point start_time;
+    std::chrono::duration<double, std::milli> accumulated_time = std::chrono::milliseconds(0);
+    std::string name;
+
+public:
+    // Constructor to set the name of the timer for identification
+    Timer(const std::string& name) : name(name) {}
+
+    // Start the timer
+    void start() {
+        start_time = std::chrono::high_resolution_clock::now();
+    }
+
+    // Stop the timer and accumulate the time elapsed since the last start
+    void stop() {
+        auto end_time = std::chrono::high_resolution_clock::now();
+        accumulated_time += end_time - start_time;
+    }
+
+    // Reset the timer to zero
+    void reset() {
+        accumulated_time = std::chrono::milliseconds(0);
+    }
+
+    // Get the total accumulated time
+    double getAccumulatedTime() const {
+        return accumulated_time.count();
+    }
+
+    // Display the accumulated time
+    void printAccumulatedTime() const {
+        std::cout << name << " total time: " << getAccumulatedTime() << " milliseconds" << std::endl;
+    }
+
+    // Display the accumulated time adjusted for a baseline accumulated time
+    void printAccumulatedTimeAdjustedForBaseline(double baseline) const {
+        std::cout << name << " total time: " << (getAccumulatedTime() - baseline) << " milliseconds" << std::endl;
+    }
+
+    void printTimePerInvocation(unsigned int sample_count) const {
+        std::cout << name << " time per invocation: " << getAccumulatedTime()*1.0e6f/float(sample_count) << " nanoseconds" << std::endl;
+    }
+
+};
+
+vec3 calculate_wi_local(float theta_i) {
+    return vec3(std::sin(radians(theta_i)), 0.0f, std::cos(radians(theta_i)));
+}
+
+vec3 calculate_wo_local(float theta_o, float phi) {
+    return vec3(std::sin(radians(theta_o)) * std::cos(radians(phi)),
+                std::sin(radians(theta_o)) * std::sin(radians(phi)),
+                std::cos(radians(theta_o)));
+}
+
[…]

### 2025-12-19 code/brdf_plotting.py

@@ -1,0 +1,123 @@
+# Prerequisites:
+#
+#     python3 -m pip install matplotlib pandas
+#
+# Running this script:
+#
+#     python3 brdf_plotting.py
+#
+# Viewing the output:
+#
+#     open brdf_plots.pdf
+#
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import glob
+import colorsys
+
+leg_locs = ['lower left', 'lower right', 'upper left',
+            'lower left', 'lower right', 'upper left',
+            'lower left', 'lower right', 'upper left']
+
+
+def plot_brdf_data():
+
+    # Path pattern to match CSV files
+    files = glob.glob('brdf_theta_i_*_phi_*.csv')
+    files.sort()
+
+    # Prepare a 3x3 subplot
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(10, 10))
+    axes = axes.flatten()
+
+    # Helper function to determine the subplot index based on theta_i and phi values
+    def get_index(theta_i, phi):
+        theta_index = {'45': 0, '80': 1}
+        phi_index   = {'0': 0, '45': 2}
+        return phi_index[phi] + theta_index[theta_i]
+
+    for file in files:
+        data = pd.read_csv(file)
+
+        # Extract theta_i and phi from the filename
+        parts = file.replace('.csv', '').split('_')
+        theta_i = parts[3]
+        phi = parts[5]
+        roughness = float(parts[7])  # roughness
+
+        print(f"Processing file: {file} with theta_i: {theta_i}, phi: {phi}, roughness: {roughness}")
+
+        # Convert theta_i to radians and calculate cos(theta_i)
+        theta_i_rad = np.radians(float(theta_i))
+        cos_theta_i = np.cos(theta_i_rad)
+
+        # Adjust the BRDF values by multiplying by cos(theta_i)
+        #data['Lambert'] *= cos_theta_i
+        #data['ON'] *= cos_theta_i
+        #data['QON'] *= cos_theta_i
+        #data['footnoteQON'] *= cos_theta_i
+        #data['fujiiQON'] *= cos_theta_i
+        #data['FON'] *= cos_theta_i
+        #data['EON'] *= cos_theta_i
+        #data['GGX'] *= cos_theta_i
+        #data['MRM'] *= cos_theta_i
+
+        # Calculate the correct subplot index based on extracted theta_i and phi
+        index = get_index(theta_i, phi)
+
+        # Set custom x-axis ticks
[…]

### 2025-12-19 code/brdf_profiler.cpp

@@ -1,0 +1,187 @@
+/*
+Prerequisites:
+
+- Clone the glm repository from https://github.com/g-truc/glm
+- Set the path to the glm repository in the include directive below
+
+General compilation and running commands:
+
+g++ -std=c++17 -O3 -o brdf_profiler brdf_profiler.cpp -I/path/to/glm
+./brdf_profiler
+
+Compilation and running commands for Peter's setup:
+
+cd ~/repos/EON_jcgt_paper/code
+g++ -std=c++17 -O3 -o brdf_profiler brdf_profiler.cpp -I/Users/pkutz/repos/glm/
+./brdf_profiler
+
+Compilation and running script for Peter's setup (modify for your own setup):
+
+./profile_brdfs.sh
+*/
+
+#include "brdf_common.h"
+
+#include <algorithm>
+#include <chrono>
+#include <random>
+
+class BRDFProfiler {
+public:
+    // This function profiles all of the BRDF functions.
+    // Details:
+    //     - It runs on a single thread.
+    //     - It profiles only the BRDF evaluation (no importance sampling).
+    //     - It interleaves different BRDFs to avoid bias due to system load varying over time.
+    //     - It varies all BRDF inputs to make sure different performance characteristics are covered.
+    //     - It calculates random directions outside the timed loop to avoid direction sampling overhead. 
+    //     - It compares to a no-op baseline to compensate for timing and looping overhead.
+    //     - It should be run with realistic compiler optimizations enabled.
+    //     - It takes various measures to avoid code disappearing due to compiler optimizations (important).
+    void profileBRDFs() {
+        // Create timer instances for each BRDF.
+        Timer noOpTimer("No-op"); // Baseline reference.
+        Timer lambertTimer("Lambert");
+        Timer fullOnTimer("Full ON");
+        Timer qonTimer("QON");
+        Timer footnoteQonTimer("Footnote QON");
+        Timer fujiiQonTimer("Fujii QON");
+        Timer fonTimer("FON");
+        Timer eonExactTimer("EON (exact)");
+        Timer eonApproxTimer("EON (approx)");
+
+        // Accumulate and print the total BRDF value to avoid optimization.
+        vec3 totalBrdfValue = vec3(0.0f);
+
[…]

### 2025-12-19 code/plot_brdfs.sh

@@ -1,0 +1,1 @@
+g++ -std=c++17 -O3 -o brdf_calculator brdf_calculator.cpp -I/Users/pkutz/repos/glm/ && ./brdf_calculator && python3 brdf_plotting.py && open brdf_plots.pdf

### 2025-12-19 code/profile_brdfs.sh

@@ -1,0 +1,1 @@
+g++ -std=c++17 -O3 -o brdf_profiler brdf_profiler.cpp -I/Users/pkutz/repos/glm/ && ./brdf_profiler > brdf_profiling_results.txt

### 2025-12-19 code/sampling_profiler.cpp

@@ -1,0 +1,324 @@
+/*
+Prerequisites:
+
+- Clone the glm repository from https://github.com/g-truc/glm
+- Set the path to the glm repository in the include directive below
+
+General compilation and running commands:
+
+g++ -std=c++17 -O3 -o brdf_profiler brdf_profiler.cpp -I/path/to/glm
+./brdf_profiler
+
+Compilation and running commands for Peter's setup:
+
+cd ~/repos/EON_jcgt_paper/code
+g++ -std=c++17 -O3 -o brdf_profiler brdf_profiler.cpp -I/Users/pkutz/repos/glm/
+./brdf_profiler
+
+Compilation and running script for Peter's setup (modify for your own setup):
+
+./profile_brdfs.sh
+*/
+
+#include "brdf_common.h"
+
+#include <algorithm>
+#include <chrono>
+#include <random>
+
+
+class BRDFProfiler {
+public:
+    // This function profiles all of the BRDF functions.
+    // Details:
+    //     - It runs on a single thread.
+    //     - It profiles only the BRDF evaluation (no importance sampling).
+    //     - It interleaves different BRDFs to avoid bias due to system load varying over time.
+    //     - It varies all BRDF inputs to make sure different performance characteristics are covered.
+    //     - It calculates random directions outside the timed loop to avoid direction sampling overhead.
+    //     - It compares to a no-op baseline to compensate for timing and looping overhead.
+    //     - It should be run with realistic compiler optimizations enabled.
+    //     - It takes various measures to avoid code disappearing due to compiler optimizations (important).
+    void profileBRDFs() {
+        // Create timer instances for each BRDF.
+        Timer noOpTimer("No-op"); // Baseline reference.
+
+        Timer lambertTimer("Lambert eval");
+        Timer lambertCosSampleTimer("Lambert sample (cos)");
+
+        Timer fullOnTimer("Full ON eval");
+        Timer fullOnCosampleTimer("Full ON sample (cos)");
+
+        Timer qonTimer("QON eval");
+        Timer qonCosSampleTimer("QON sample (cos)");
+
+        Timer footnoteQonTimer("Footnote QON eval");
+        Timer footnoteQonCosSampleTimer("Footnote QON sample (cos)");
+
[…]

### 2025-12-19 article.tex
Fix placements.

@@
 \usepackage{natbib}
+\usepackage[font={small,it}]{caption}
 
@@
   \vspace*{-0.5cm}
-    \hfill
-    \includegraphics[width=0.25\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
-    \includegraphics[width=0.25\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
-    \includegraphics[width=0.25\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
+    \includegraphics[width=0.3\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
+    \includegraphics[width=0.3\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
+    \includegraphics[width=0.3\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
     \vspace{0.06cm}
-    \includegraphics[width=0.25\linewidth]{figures/cone/retro_cone_cropped_label.png}
-    \includegraphics[width=0.25\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
-    \includegraphics[width=0.25\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
-    \hfill
+    \includegraphics[width=0.3\linewidth]{figures/cone/retro_cone_cropped_label.png}
+    \includegraphics[width=0.3\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
+    \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
   \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
@@
 
-Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward:
+Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):
 
@@
   As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
-\item Pseudo-code is provided in Listing~\ref{lst:implementation}.
 \end{itemize}
@@
     language=GLSL,
-    caption=Pseudo code to implement the retro-reflective microfacet model using existing regular microfacet BSDF implementation,
+    caption=Pseudo code to implement the retro-reflective microfacet model using an existing regular microfacet BSDF implementation,
     label=lst:implementation,
@@
 
-\begin{figure}[!b]
+\paragraph*{Combination of BRDF and BTDF}
+
+The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
+When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+
+Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
+Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
+
+\begin{figure}[!ht]
     \centering
@@
     \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
+  \captionsetup{font=small}
   \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction.
-The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).}
+The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).
+\vspace{-0.5cm}}
   \label{fig:conductor_renders}
@@
 
-\paragraph*{Combination of BRDF and BTDF}
-
-The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
-When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
-
-Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
-Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
-\begin{figure}[!tb]
+\begin{figure}[!hb]
     \centering

### 2025-12-19 article.tex
Fixing a typo in section 2
Moving one figure to tikz

@@
 \usepackage[font={small,it}]{caption}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+
+\usepackage[export]{adjustbox}
 
@@
 In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
-In this work, we show that we can use the back vector~\cite{BelcourBackvector} used to model retro-reflection as a microfacet selector.
+In this work, we show that we can use the back vector~\cite{BelcourBackvector} as a microfacet selector to model retro-reflection.
 Before detailing the impact of such a change, we review some properties of microfacet models.
@@
 \end{equation}
-This is due to the fact that we have to differentiate with respect to the light direction (which is not altered by the back vector).
+This is due to the fact that we have to differentiate with respect to the light direction which is not altered by the back vector.
 
@@
   \centering
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
-    \vspace{0.06cm}
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
-    \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
+  \begin{tikzpicture}[font=\footnotesize]
+    \node[inner sep=1pt] (Class1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }};
+    \node[inner sep=1pt,right = 0cm of Class1] (MRM1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }};
+    \node[inner sep=1pt,right = 0cm of MRM1] (Mixed1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png }};
+
+    \node[inner sep=1pt,below = 0cm of Class1] (Class15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }};
+    \node[inner sep=1pt,right = 0cm of Class15] (MRM15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png }};
+    \node[inner sep=1pt,right = 0cm of MRM15] (Mixed15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }};
+
+    \node[inner sep=1pt,above = 2pt of Class1] { \textbf{Classical Microfacet} };
+    \node[inner sep=1pt,above = 2pt of MRM1] { \textbf{Full retro-reflective MRM} };
+    \node[inner sep=1pt,above = 11.5pt of Mixed1] { \textbf{Mixed MRM (BRDF)} };
+    \node[inner sep=1pt,above = 0.5pt of Mixed1] { \textbf{and Classical (BTDF)} };
+
+    \node[inner sep=1pt,left = 1pt of Class1, anchor=south,rotate=90] { $\mathbf{\boldsymbol{\eta} = 1}$ };
+    \node[inner sep=1pt,left = 1pt of Class15, anchor=south,rotate=90] { $\mathbf{\boldsymbol{\eta} = 1.5}$ };
+  \end{tikzpicture}
+
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
+    % \vspace{0.06cm}
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
+    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
   \caption{Renders of a glass shaderball illuminated from the side.

### 2025-12-19 article.tex
Changing writing of G1(l,v) to G1(l,h) in exposition

@@
 \begin{equation}
-G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vecl), G_1(\vecl, \vecv) \right) \ .
+G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vech), G_1(\vecl, \vech) \right) \ .
 \end{equation}
 Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
-$G_{2}(\vecv, \vecl) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
+$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
 \begin{equation}\label{eq:visible}
-  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \,G_1(\vecv, \text{reflect}(\vecv, \vech)) \vdot{\vecv}{\vech} d\vech.
+  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \,G_1(\vecv, \vech) \vdot{\vecv}{\vech} d\vech.
 \end{equation}
@@
 \begin{equation}
-  G_1(\vecv, \vecl) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
+  G_1(\vecv, \vech) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
 \end{equation}
@@
 \begin{equation}
-  G_1(\vecv, \vecl) = G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
+  G_1(\vecv, \vech) = G_{1,\text{vc}, \vech}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
   \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
@@
   \begin{split}
-  G_{1,\text{Smith}}(\vecv) &=
+  G_{1,\text{Smith}}(\vecv, \vech) &=
   \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
   &= \frac{\avdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecvv}: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =
-  G_{1,\text{Smith}}(\vecvv)
+  G_{1,\text{Smith}}(\vecvv, \vech)
   \end{split}
@@
 \end{equation}
-Further, using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get
+Getting back to the geometry term, we first note that we can rewrite $G_1(\vecv, \vech)$ as $G_1(\vecv, \vech(\vecv, \vecl))$.
+For ease, we write $G_1(\vecv, \vecl) = G_1(\vecv, \vech)$, and $G_1(\vecl, \vecv) = G_1(\vecl, \vech)$.
+It follows that using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get
 \begin{equation}

### 2025-12-19 article.tex
change missing notation to h and use words instead of =

@@
 \paragraph*{Shadowing-Masking.}
-The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vecl)$ for view direction and
-$G_1(\vecl, \vecv)$ for light direction:
+The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vech)$ for view direction and
+$G_1(\vecl, \vech)$ for light direction:
 \begin{equation}
@@
 Getting back to the geometry term, we first note that we can rewrite $G_1(\vecv, \vech)$ as $G_1(\vecv, \vech(\vecv, \vecl))$.
-For ease, we write $G_1(\vecv, \vecl) = G_1(\vecv, \vech)$, and $G_1(\vecl, \vecv) = G_1(\vecl, \vech)$.
+For ease, we change the notation to $G_1(\vecv, \vecl)$ instead of $G_1(\vecv, \vech)$, and $G_1(\vecl, \vecv)$ instead of $G_1(\vecl, \vech)$.
 It follows that using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get

### 2025-12-19 article.tex
use Vp instead of V in source code

@@
     vec3 Vp = -V + 2.0 * dot(V, N) * N;
-    return sample_bsdf(V, L, N, params);
+    return sample_bsdf(Vp, L, N, params);
 }
@@
     vec3 Vp = -V + 2.0 * dot(V, N) * N;
-    return eval_bsdf(V, L, N, params);
+    return eval_bsdf(Vp, L, N, params);
 }

### 2025-12-19 article.tex
Fixing math G2
Small fixes.

@@
 % Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
-\accepted{yyyy-mm-dd}{yyyy-mm-dd}{yyyy-mm-dd}{Editor Name}{volume}{issue}{1}{6}{yyyy}
+%\accepted{yyyy-mm-dd}{yyyy-mm-dd}{yyyy-mm-dd}{Editor Name}{volume}{issue}{1}{6}{yyyy}
 \seturl{http://jcgt.org/published/vol/issue/num/}
@@
 
-\newcommand{\todo}[1]{{\color{red}\textbf{[TODO:} #1\textbf{]}}}
 \newcommand{\laurent}[1]{{\color{blue}\textbf{[Laurent:} #1\textbf{]}}}
@@
   \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
-The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak.}
+The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak in the MRM model.}
    \vspace{-0.5cm}
@@
 \maketitle
-\thispagestyle{firstpagestyle}
+%\thispagestyle{firstpagestyle}
 
@@
 
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and follows measured observations, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
-We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets these requirements.
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
+We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets all these requirements.
 It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
@@
 However, their model only provides pure retro-reflection which would lack artistic control.
-They later extented this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
+They later extended this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
 However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
@@
 \begin{equation}
-G_{2}(\vecv, \vecl) = G_{2}\left(G_1(\vecv, \vech), G_1(\vecl, \vech) \right) \ .
+G_{2}(\vecv, \vecl) = G_1(\vecv, \vech) \times G_1(\vecl, \vech)\ .
 \end{equation}
@@
 For v-cavities masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
-%todo{Laurent: I think we could extract the shadowing from Matthias derivation here.
+%\todo{\tiny Laurent: I think we could extract the shadowing from Matthias derivation here.
 The other way to justify the derivation of the full model is to highlight that Smith's shadowing/masking does not depend on the microfacet (appart from the sideness).

### 2025-12-21 article.tex
Add data comparison.

@@
 
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=0.99\linewidth]{figures/brdf_plots.pdf}
-  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
-The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$.
-The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
-  \label{fig:brdf_plots}
-\end{figure}
-
 \paragraph*{Microfacet reflection term.}
@@
 we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
+
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/brdf_plots.pdf}
+  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
+The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$.
+The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
+  \label{fig:brdf_plots}
+\end{figure}
 
@@
 
-\noindent
+%\noindent
 \begin{lstlisting}[
@@
 
-\paragraph*{Incorporation in a standard microfacet model}
+\paragraph*{Comparison to measured data}
 
-Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retro-reflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.
+Figure~\ref{fig:data_fit} shows a fit of the retro-reflective MRM model (in red) to the measured BRDF of retro-reflective tape (in blue) from \cite{BelcourBackvector}.
+This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
+The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel with $F_0=0.19$ and $F_{82} = 0$.
 
-\paragraph*{Combination of BRDF and BTDF}
+\begin{figure}[!b]
+  \begin{center}
+    \includegraphics[width=0.95\textwidth]{figures/data_fit.png}
+  \end{center}
+  \caption{Measured BRDF of retro-reflective tape (blue) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red).}
+  \label{fig:data_fit}
+\end{figure}
 
-The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
-When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+\newpage
 
-Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
-Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
-
-\begin{figure}[!ht]
+\begin{figure}[!!t]
     \centering
@@
 
-\begin{figure}[!hb]
+\begin{figure}[!b]
   \centering
@@
   \end{tikzpicture}
-
     % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
@@
 \end{figure}
+
+\newpage
+
+\paragraph*{Incorporation in a standard microfacet model}
+
+Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retro-reflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.
+
+\paragraph*{Combination of BRDF and BTDF}
+
+The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
+When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+
+Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
+Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
 

### 2025-12-21 code/brdf_calculator_mrm.cpp

@@ -1,0 +1,88 @@
+/*
+Prerequisites:
+
+- Clone the glm repository from https://github.com/g-truc/glm
+- Set the path to the glm repository in the include directive below
+
+General compilation and running commands:
+
+g++ -std=c++17 -O3 -o brdf_calculator brdf_calculator.cpp -I/path/to/glm
+./brdf_calculator
+
+Compilation and running commands for Peter's setup:
+
+cd ~/repos/EON_jcgt_paper/code
+g++ -std=c++17 -O3 -o brdf_calculator brdf_calculator.cpp -I/Users/pkutz/repos/glm/
+./brdf_calculator
+
+Compilation and running script for Peter's setup (modify for your own setup):
+
+./plot_brdfs.sh
+*/
+
+#include "brdf_common.h"
+
+#include <fstream>
+#include <string>
+#include <vector>
+
+
+float deg_to_rad(float degrees)
+{
+    return degrees * (3.14159265358979323846f / 180.0f);
+}
+
+class BRDFCalculator {
+public:
+    void computeAndOutputBRDFData() {
+        vec3 rho = vec3(1.f);  // Albedo.
+
+        // ############################# MRM model parameters #############################
+        std::vector<float> roughnesses = {0.23f}; // Roughness values.
+        // ############################# MRM model parameters #############################
+
+        for (float roughness : roughnesses)
+        {
+            std::vector<float> thetas_i = {15.0f, 30.0f, 60.f}; // Incident zenith angles in degrees.
+            std::vector<float> phis = {0.0f};     // Relative azimuth angles in degrees.
+
+            // Loop through each combination of theta_i and phi to generate separate CSV files.
+            for (float theta_i : thetas_i) {
+
+                for (float phi_o : phis) {
+                    std::string filename = "brdf_mrm_theta_i_" + std::to_string((int)theta_i) + "_phi_" + std::to_string((int)phi_o) + "_roughness_" + std::to_string((roughness)) + ".csv";
+                    std::ofstream file(filename);
+
+                    // Write headers to the CSV file.
+                    file << "theta_o,MRM\n";
+
+                    float angle_tol = 1.0e-4f;
+                    // Compute BRDF values over the range of theta_o.
+                    for (float theta_o = -90.0f+angle_tol; theta_o <= 90.0f-angle_tol; theta_o += 1.0f) {
+                        vec3 wi_local = calculate_wi_local(theta_i);
+                        vec3 wo_local = calculate_wo_local(theta_o, phi_o);
+
[…]

### 2025-12-21 code/brdf_plotting.py

@@ -99,8 +99,8 @@ def plot_brdf_data():
         axes[index].plot(data['theta_o'], data['MRM'],     label='MRM'     if show_legend else None, color='green',  linewidth=1.0, linestyle=linestyle)
 
         # Set the title for each subplot
-        axes[index].set_title(f"$\\theta_i$: ${theta_i}^\circ$, $\\phi_o$: ${phi}^\circ$, $r$: ${roughness}$", fontsize=12)
-        axes[index].set_xlabel(r"$\theta_o$", fontsize=16, fontweight='bold')
+        axes[index].set_title(f"$\\theta_i: {theta_i}^\\circ$, $\\phi_o: {phi}^\\circ$, $r: {roughness}$", fontsize=12)
+        axes[index].set_xlabel(f"$\\theta_o$", fontsize=16, fontweight='bold')
         axes[index].set_ylabel('BRDF value', fontsize=16)
         axes[index].set_xlim(-90, 90)
 

### 2025-12-21 code/brdf_plotting_with_data.py

@@ -1,0 +1,143 @@
+# Prerequisites:
+#
+#     python3 -m pip install matplotlib pandas
+#
+# Running this script:
+#
+#     python3 brdf_plotting.py
+#
+# Viewing the output:
+#
+#     open brdf_plots.pdf
+#
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import glob
+import colorsys
+import math
+
+#################################
+
+def rad_to_deg(rad):
+    return rad * 180.0 / math.pi
+
+# Load space-separated BRDF data from files and plot them
+
+# Load data, skipping comment lines
+data = np.loadtxt('../data/retro-3M-jaune.dat', comments='#')
+
+# Extract columns
+theta_L = data[:, 0]      # First column
+theta_V = data[:, 1]      # Second column
+phi_diff = data[:, 2]     # Third column
+brdf_values_times_cosine = data[:, 3]  # Fourth column
+
+# Quantize values to differ at most by 1.0e-10
+quantization_precision = 1.0e-10
+theta_L = np.round(theta_L / quantization_precision) * quantization_precision
+theta_V = np.round(theta_V / quantization_precision) * quantization_precision
+phi_diff = np.round(phi_diff / quantization_precision) * quantization_precision
+brdf_values = np.round(brdf_values_times_cosine / quantization_precision) * quantization_precision
+
+# Filter out rows where abs(phi_diff) > 0.0001
+mask = np.abs(phi_diff) <= 0.0001
+theta_L = theta_L[mask]
+theta_V = theta_V[mask]
+phi_diff = phi_diff[mask]
+brdf_values_times_cosine = brdf_values_times_cosine[mask]
+
+
+
+# Create a dictionary mapping theta_L to list of (theta_V, brdf_value) pairs
+brdf_dict_simple = {}
+for tL, tV, brdf_times_cos in zip(theta_L, theta_V, brdf_values_times_cosine):
+    if tL not in brdf_dict_simple:
+        brdf_dict_simple[tL] = []
+    print('tL:', tL, 'cos(tL):', np.cos(tL), 'brdf_times_cos:', brdf_times_cos)
+    brdf_dict_simple[tL].append((tV, brdf_times_cos / np.cos(tL)))  # Store BRDF value (not multiplied by cosine)
+
+
[…]

### 2025-12-21 article.tex
Mirror diagram improvements. Some text fixes/improvements.

@@
 
-\begin{figure}
+\begin{figure}[!b]
   \begin{center}
@@
       \hline
-      $\vecn$ & surface normal, with the convention that it points to
-      the medium with lower index of refraction in cases dealing with $\vecv$ and $\vecl$ pointing to different sides\\
+      $\vecn$ & surface macronormal, with the convention in the refractive case that it points into the medium with a lower IOR \\
       \hline
-      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the normal) \\
+      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the macronormal) \\
       \hline
@@
       \hline
-      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to (always points to medium with lower index of refraction)\\
+      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to\\
       \hline
@@
 where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
-The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
-Similarly, using the half-vector of refraction as selector, yields the corresponding BTDF
+The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
+Similarly, using the half-vector of refraction $\vecht$ as selector yields the corresponding BTDF:
 \begin{equation}
@@
 In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
-In this work, we show that we can use the back vector~\cite{BelcourBackvector} as a microfacet selector to model retro-reflection.
-Before detailing the impact of such a change, we review some properties of microfacet models.
+In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retro-reflection.
+Before detailing the impact of such a change, we review some standard properties of microfacet models.
 
 \paragraph*{Shadowing-Masking.}
-The shadowing-masking function combines the visibility masking function $G_1(\vecv, \vech)$ for view direction and
-$G_1(\vecl, \vech)$ for light direction:
+
+The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function
+$G_1(\vecl, \vech)$ for the light direction.
+In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
+non-separable shadowing-masking functions.
+Our derivation holds for any valid $G_2$ function, but we focus here on the separable case for simplicity.}
 \begin{equation}
-G_{2}(\vecv, \vecl) = G_1(\vecv, \vech) \times G_1(\vecl, \vech)\ .
+G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \times G_1(\vecl, \vech)\ .
 \end{equation}
@@
 \end{equation}
-where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function
+where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
 \begin{equation}
@@
   \begin{center}
-    \includegraphics[width=0.4\textwidth]{figures/backvector.pdf}
-    \includegraphics[width=0.4\textwidth]{figures/backvector_refract.pdf}
+    \includegraphics[width=0.3\textwidth]{figures/backvector.pdf}
+    \includegraphics[width=0.3\textwidth]{figures/backvector_refract.pdf}
   \end{center}
-  \caption{Geometry of the back-vector for reflection and refraction, illustrated in two dimensions.}
+  \caption{Geometry of the back-vector for reflection and refraction, illustrated in two dimensions.
+The view vector $\vecv$ is reflected in a ``virtual mirror'' to obtain $\vecvv$, which is then used to compute the half-vector with the light vector $\vecl$.
+The resulting back-vector ($\vecb_{r}$ for reflection and $\vecb_{t}$ for refraction) is used as microfacet normal selector.}
   \label{fig:backvector}
@@
 
-As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual'' plane aligned with $\vecv$ and perpendicular to the surface.
-Intuitively, this is equivalent to observing the microfacet reflection through a virtual mirror, thus turning glossy forward-reflection into glossy retro-reflection.
+As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a virtual plane aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$.
+Intuitively, this is equivalent to observing the microfacet reflection through a ``virtual mirror'', thus turning glossy forward-reflection into glossy retro-reflection.
 While there is no immediate physical foundation to this construction, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
@@
 There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
-As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector instead.
+As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead.
 From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
@@
 \end{equation}
+Thus for both reflection and refraction, the effect on the microsurface BSDF of using the back-vector as selector is to replace $\vecv$ by $\vecvv$ in the Fresnel term and in the denominators.
 
 \paragraph*{Shadowing \& Masking.}
-We simply apply Smith or v-cavities masking to the mirrored microfacet distribution.
+We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution.
 With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
-For v-cavities masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
+For V-cavity masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
 %\todo{\tiny Laurent: I think we could extract the shadowing from Matthias derivation here.
@@
 
+We briefly prove here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills symmetry and energy conservation requirements.
+
 In analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
@@
   &= -(\vecll + \vecv) + 2 \vdot{(\vecll + \vecv)}{\vecn} \vecn =
-  \text{reflect}(\vecll + \vecv, \vecn) \ .
+  \text{reflect}(\vecll + \vecv, \vecn) \ ,
   \end{split}
@@
   &= -(\etal\vecll + \etav\vecv) + 2 \vdot{(\etal\vecll + \etav\vecv)}{\vecn} \vecn =
-  \text{reflect}(\etal\vecll + \etav\vecv, \vecn) \ .
+  \text{reflect}(\etal\vecll + \etav\vecv, \vecn)
   \end{split}
@@
   f_r(\vecvv, \vecl) &=
-  \frac{D(\vech(\vecvv, \vecl)) G_{2}(G_1(\vecvv, \vecl), G_1(\vecl, \vecvv)) F(\vecvv, \vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
+  \frac{D(\vech(\vecvv, \vecl)) \, G_{2}(\vecvv, \vecl) \, F\left(\vecvv, \vech(\vecvv, \vecl)\right)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
   &=
-  \frac{D(\vech(\vecll, \vecv)) G_{2}(G_1(\vecv, \vecll), G_1(\vecll, \vecv)) F(\vecll, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} }\\
+  \frac{D(\vech(\vecll, \vecv)) \, G_{2}(\vecll, \vecv) \, F\left(\vecll, \vech(\vecll,\vecv)\right)}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} }\\
   &=
@@
   \begin{center}
-    \includegraphics[width=0.95\textwidth]{figures/data_fit.png}
+    \includegraphics[width=0.99\textwidth]{figures/brdf_theta_L_plots.pdf}
   \end{center}
-  \caption{Measured BRDF of retro-reflective tape (blue) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red).}
+  \caption{Measured BRDF of retro-reflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
   \label{fig:data_fit}

### 2025-12-22 article.tex
Simplify reciprocity discussion.

@@
 \newcommand{\etay}{\eta_{\mathbf{y}}}
-\newcommand{\vdot}[2]{\left\langle #1, #2 \right\rangle}
+\newcommand{\vdot}[2]{#1 \cdot #2}
 %\newcommand{\vdot}[2]{#1\cdot#2}
 \newcommand{\avdot}[2]{\lvert\vdot{#1}{#2}\rvert}
+\newcommand{\cldot}[2]{\langle #1 , #2 \rangle}
+
+\newcommand{\diff}{\, \mathrm{d}}
 
@@
 
-\section{Microfacet BSDF Models}
+\section{Back-vector Modification of Microfacet Models}
 \label{sec:microfacets}
 
-\begin{figure}[!b]
-  \begin{center}
-    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
-%    \begin{tabular} {|l|l|}
-      \hline
-      $\vecv$ & normalized vector to viewer \\
-      \hline
-      $\vecl$ & normalized vector to light \\
-      \hline
-      $\vecn$ & surface macronormal, with the convention in the refractive case that it points into the medium with a lower IOR \\
-      \hline
-      $\mathbf{t_1}, \mathbf{t_2}$ & tangent and bitangent (forming an orthonormal coordinate frame with the macronormal) \\
-      \hline
-      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
-      \hline
-      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, where $\etav$ and $\etal$ are the refractive indices of the media $\vecv$ and $\vecl$ point to\\
-      \hline
-      $\vdot{\vecx}{\vecy}$ & dot product of $\vecx$ and $\vecy$ \\
-      \hline
-      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e.
-$\text{reflect}(\mathbf{x}, \mathbf{m}) = -\mathbf{x} + 2 \vdot{\mathbf{x}}{\mathbf{m}} \mathbf{m}$ \\
+%\begin{figure}[!b]
+%  \begin{center}
+%    \begin{tabular} {|p{0.2\textwidth}|p{0.7\textwidth}|}
+%%    \begin{tabular} {|l|l|}
 %      \hline
-%      $\Omega$ & hemisphere of directions around the normal\\
-      \hline
-      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{
-    Notation used.
-  }
-  \label{fig:notation}
-\end{figure}
+%      $\vecv$ & normalized vector to viewer \\
+%      \hline
+%      $\vecl$ & normalized vector to light \\
+%      \hline
+%      $\vecn$ & surface macronormal, with the convention in the refractive case that it points into the medium with a lower IOR \\
+%      \hline
+%      $\mathbf{t_1}, \mathbf{t_2}$ & tangent, bitangent (forming an orthonormal frame with $\vecn$) \\
+%      \hline
+%      $\vechr$ & half vector of reflection, $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ \\
+%      \hline
+%      $\vecht$ & half vector of refraction, $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$, %where $\etav$, $\etal$ are the IORs of the media $\vecv$, $\vecl$ point to\\
+%      \hline
+%      $\cldot{\vecx}{\vecy}$ & clamped dot product of $\vecx$ and $\vecy$, i.e. $\max(0, \vecx \cdot \vecy)$ \\
+%      \hline
+%      $\chi^+(a)$ & Heaviside step function, i.e. $1$ if $a > 0$ and $0$ otherwise \\
+%      \hline
+%      $\text{reflect}(\mathbf{x}, \mathbf{m})$ & reflection of vector $\mathbf{x}$ on surface with normal $\mathbf{m}$, i.e.
+$\text{reflect}(\mathbf{x}, \mathbf%{m}) = -\mathbf{x} + 2 (\vdot{\mathbf{x}}{\mathbf{m}})\mathbf{m}$ \\
+%%      \hline
+%%      $\Omega$ & hemisphere of directions around the normal\\
+%      \hline
+%      $\Omega, \Omega_{\vecx}$ & hemisphere of directions around the normal, hemisphere of directions around the normal pointing to the side $\vecx$ points to\\
+%      \hline
+%    \end{tabular}
+%  \end{center}
+%  \caption{
+%    Notation used.
+%  }
+%  \label{fig:notation}
+%\end{figure}
 
-Using the notation of Figure \ref{fig:notation}, a microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
+A microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
 \begin{equation}
-  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl) \, F(\vechr, \vecv)}{4 \vdot{\vecv}{\vecn} \vdot{\vecl}{\vecn}}
+  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl,\vechr) \, F(\vecv, \vechr)}{4 \avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn}}
   \label{eq:microfacet_brdf}
 \end{equation}
-where $D$ is the distribution of microfacets, normalized to fulfill $\int_\Omega D(\vech) \vdot{\vecn}{\vech} d\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
-The half-vector $\vechr$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
-Similarly, using the half-vector of refraction $\vecht$ as selector yields the corresponding BTDF:
+where $D$ is the distribution of micronormals (NDF), normalized to fulfill $\int_\Omega D(\vech) (\vdot{\vecn}{\vech}) \diff\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
+The half-vector $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
+Similarly, using the half-vector of refraction $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$ (where $\etav$, $\etal$ are the IORs of the media $\vecv$, $\vecl$ point to) as selector yields the corresponding BTDF:
 \begin{equation}
   f_t(\vecv, \vecl) =
-  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\etav^2 D(\vecht) \, G_{2}(\vecv,\vecl) \, \left(1 - F(\vecht, \vecv) \right)}
+  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\, \etav^2 \, D(\vecht) \, G_{2}(\vecv,\vecl,\vecht) \, \bigl(1 - F(\vecv, \vecht) \bigr)}
        {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
@@
 In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retro-reflection.
-Before detailing the impact of such a change, we review some standard properties of microfacet models.
+%Before detailing the impact of such a change, we review some standard properties of microfacet models.
 
-\paragraph*{Shadowing-Masking.}
+%\paragraph*{Shadowing-Masking.}
+%
+%The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility %shadowing function
+%$G_1(\vecl, \vech)$ for the light direction.
+In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of %\citet{HeitzMicrofacet} for a discussion of separable vs.
+non-separable shadowing-masking functions.
+Our derivation holds for any valid $G_2$ function, but we %focus here on the separable case for simplicity.}
+%\begin{equation}
+%G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
+%\end{equation}
+%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
+%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
+%\begin{equation}\label{eq:visible}
+%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
+%\end{equation}
+%Popular choices for $G_1$ with this property are the Smith masking function
+%\begin{equation}
+%  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
+%\end{equation}
+%where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
+%\begin{equation}
+%  G_{1,\text{vc}}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
+%  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
+%\end{equation}
+%which is generally applicable for distributions with symmetry around the normal.
 
-The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function
-$G_1(\vecl, \vech)$ for the light direction.
-In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
-non-separable shadowing-masking functions.
-Our derivation holds for any valid $G_2$ function, but we focus here on the separable case for simplicity.}
-\begin{equation}
-G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \times G_1(\vecl, \vech)\ .
-\end{equation}
-Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
-$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
-\begin{equation}\label{eq:visible}
-  \avdot{\vecv}{\vecn} = \int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \,G_1(\vecv, \vech) \vdot{\vecv}{\vech} d\vech.
-\end{equation}
-Choices for $G_1$ with this property are the Smith masking function
-\begin{equation}
-  G_1(\vecv, \vech) = G_{1,\text{Smith}}(\vecv) = \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech},
-\end{equation}
-where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
-\begin{equation}
-  G_1(\vecv, \vech) = G_{1,\text{vc}, \vech}(\vdot{\vecv}{\vecn},\vdot{\vecn}{\vech},\vdot{\vecv}{\vech}) =
-  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
-\end{equation}
-which is generally applicable for distributions with symmetry around the normal.
-
-\section{Back-vector Modification}
-\label{sec:backvector}
-
-Instead of reflection half-vector $\vechr$, we can use the reflection back-vector $\vecbr$~\cite{BelcourBackvector} defined as
+To achieve retro-reflection, we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
 \begin{equation}
@@
 \begin{equation}
-\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecb} \right|} \ .
+\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech_{r}(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecbr} \right|} \ .
 \end{equation}
@@
 \paragraph*{Shadowing \& Masking.}
+The term $G_{2}(\vecv,\vecl,\vechr)$ in the microfacet model accounts for shadowing and masking of microfacets.
+
+The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry.
+It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction.
+In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
+non-separable shadowing-masking functions.
+Our derivation holds for any valid $G_2$ function, but we focus here on the separable case for simplicity.}
+\begin{equation}
+G_{2}(\vecv, \vecl, \vech) = G_1(\vecv, \vech) \, G_1(\vecl, \vech)\ .
+\end{equation}
+Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
+$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
+\begin{equation}\label{eq:visible}
+  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
+\end{equation}
+Popular choices for $G_1$ with this property are the Smith masking function
+\begin{equation} \label{eq:smith_masking}
+  G_{1,\text{Smith}}(\vecv, \vech) = \frac{\chi^+(\vdot{\vecv}{\vech})  \avdot{\vecv}{\vecn}}{\int_{\Omega_{\vecv}} \cldot{\vecv}{\vech} D(\vech) \diff\vech},
+\end{equation}
+where analytic expressions are available for Beckmann and GGX distributions, and the V-cavity masking function \cite{HeitzMicrofacet}
+\begin{equation} \label{eq:vc_masking}
+  G_{1,\text{VC}}(\vecv, \vech) =
+  \min\left\lbrace\frac{2\avdot{\vecv}{\vecn}\avdot{\vecn}{\vech}}{\avdot{\vecv}{\vech}}, 1\right\rbrace,
+\end{equation}
+which is generally applicable for distributions with symmetry around the normal.
+
 We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution.
@@
   \begin{cases}
-    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if} \vdot{\vecv}{\vecn}\vdot{\vecl}{\vecn} \geq 0, \\
+    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
     f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
@@
 
-\begin{figure}[tbh]
+\begin{figure}[!tb]
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/brdf_plots.pdf}
+  \includegraphics[width=0.83\linewidth]{figures/brdf_plots.pdf}
   \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
@@
 
-\section{Plausibility of the Back-vector Modification}
+We briefly show here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills symmetry and energy conservation requirements.
 
-\label{sec:plausibility}
+\section{Reciprocity of MRM}
 
-We briefly prove here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills symmetry and energy conservation requirements.
+\label{sec:reciprocity}
 
-In analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
-First we observe
-\begin{equation}\label{eq:nk}
-  \begin{split}
-    \vdot{\vecvv}{\vecn} &= \vdot{\vecv}{\vecn} \text{ and}\\
-    \vdot{\vecl}{\vecn} &= \vdot{\vecll}{\vecn}.
-  \end{split}
+We first verify that the retro-reflective microfacet model fulfills reciprocity symmetry, i.e.
+\begin{equation}
+  f_{r,\text{retro}}(\vecv, \vecl) = f_{r,\text{retro}}(\vecl, \vecv) \ .
 \end{equation}
-For the reflection back-vector we show $\vechr(\vecvv, \vecl) = \text{reflect}\left(\vechr(\vecll, \vecv), \vecn\right)$, as
+For brevity, in analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
+We will verify explicitly that the retro-reflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
+
+\paragraph*{NDF term.}
+
+The NDF term $D(\vech)$ in the microfacet model selects the microfacets oriented with normal $\vech$. Under the interchange operation, we have to verify that
 \begin{equation}
-  \begin{split}
-  \vecvv + \vecl &=
-  \left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \vecl \\
-  &= \left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \left(- \vecll + 2 \vdot{\vecll}{\vecn} \vecn \right)\\
-  &= -(\vecll + \vecv) + 2 \vdot{(\vecll + \vecv)}{\vecn} \vecn =
-  \text{reflect}(\vecll + \vecv, \vecn) \ ,
-  \end{split}
+D(\vech(\vecvv, \vecl)) = D(\vech(\vecll, \vecv))
 \end{equation}
-and therefore $\lVert \vecvv + \vecl \rVert = \lVert \vecll + \vecv \rVert$ as well.
-Similarly, for refraction we have
-\begin{equation}
-  \begin{split}
-  \etav\vecvv + \etal\vecl &=
-  \etav\left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \etal\vecl \\
-  &= \etav\left(-\vecv + 2 \vdot{\vecv}{\vecn} \vecn \right) + \etal\left(- \vecll + 2 \vdot{\vecll}{\vecn} \vecn \right)\\
-  &= -(\etal\vecll + \etav\vecv) + 2 \vdot{(\etal\vecll + \etav\vecv)}{\vecn} \vecn =
-  \text{reflect}(\etal\vecll + \etav\vecv, \vecn)
-  \end{split}
-\end{equation}
-which implies $\lVert \etav\vecvv + \etal\vecl \rVert = \lVert \etal\vecll + \etal\vecv \rVert$ and
-results in $\vecht(\vecvv, \vecl) = \text{reflect}\left(\vecht(\vecll, \vecv), \vecn\right)$.
-From that we can conclude
-\begin{equation}\label{eq:nh}
-  \begin{split}
-    \vdot{\vecn}{\vech(\vecvv, \vecl)} &=
-    \vdot{\vecn}{\vech(\vecll, \vecv)}  \ , \\
-    \vdot{\mathbf{t_1}}{\vech(\vecvv, \vecl)} &=
-    -\vdot{\mathbf{t_1}}{\vech(\vecll, \vecv)} \ , \\
-    \vdot{\mathbf{t_2}}{\vech(\vecvv, \vecl)} &=
-    -\vdot{\mathbf{t_2}}{\vech(\vecll, \vecv)}\\
-    \end{split}
-\end{equation}
-for both reflection ($\vech = \vechr$) and refraction ($\vech = \vecht$) cases.
-Since we assumed $D$ is reflection-symmetric around the normal, i.e. it takes the form
+for both reflection and refraction cases.
+Since $\vech(\vecvv, \vecl)$ is the reflection of $\vech(\vecll, \vecv)$ around the normal, this holds true if $D$ is reflection-symmetric around the normal, i.e.
+it takes the form
 \begin{equation}
@@
 \end{equation}
-we can thus infer that using it with a back-vector obeys symmetry.
-Further, the reflection symmetry of $D$ along with Equation \eqref{eq:nk} gives
+which we as noted above is a requirement for our model, but true of most commonly used microfacet distributions (including for example anisotropic GGX).
+
+\paragraph*{Shadowing \& Masking.}
+
+The shadowing-masking function $G_{2}(\vecv,\vecl,\vech)$ in the microfacet model accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry.
+We have to verify that
+\begin{equation} \label{eq:shadowing_symmetry}
+G_{2}(\vecvv,\vecl,\vech(\vecvv, \vecl)) = G_{2}(\vecll,\vecv,\vech(\vecll, \vecv)) \ .
+\end{equation}
+First, V-cavity masking manifestly satisfies the following symmetry under reflection of both the view and light directions \emph{and} the selected microfacet normal:
 \begin{equation}
-  \begin{split}
-  G_{1,\text{Smith}}(\vecv, \vech) &=
-  \frac{\avdot{\vecv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecv}: \vdot{\vecv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecv}{\vech} d\vech} \\
-  &= \frac{\avdot{\vecvv}{\vecn}}{\int_{\lbrace\vech \in \Omega_{\vecvv}: \vdot{\vecvv}{\vech} \geq 0\rbrace} D(\vech) \vdot{\vecvv}{\vech} d\vech} =
-  G_{1,\text{Smith}}(\vecvv, \vech)
-  \end{split}
+  G_2(\vecvv, \vecl, \vech(\vecvv, \vecl)) = G_2(\vecv, \vecll, \vech(\vecv, \vecll)) \ .
 \end{equation}
-for Smith-masking. Next we show
+as can be verified with Equation~\eqref{eq:vc_masking}.
+In the case of Smith masking, this also holds provided that $D$ is reflection-symmetric around the normal (which we assumed), as follows from Equation~(\eqref{eq:smith_masking}.
+Then it follows from standard reciprocity that
 \begin{equation}
-  \begin{split}
-%  \vdot{\vecx}{\vecy} = -\vdot{\vecxx}{\vecy} + 2 \vdot{\vecxx}{\vecn}\vdot{\vecy}{\vecn}
-%  = \vdot{\vecxx}{-\vecyy + 2 \vdot{\vecyy}{\vecn}\vecn} = \vdot{\vecxx}{\vecyy}
-    \vdot{\vecxx}{\vecyy} &= \vdot{-\vecx}{\vecyy} + 2 \vdot{\vecx}{\vecn}\vdot{\vecn}{\vecyy}\\
-  &= \vdot{-\vecx}{-\vecy} + \vdot{-\vecx}{2 \vdot{\vecy}{\vecn}\vecn} + 2 \vdot{\vecx}{\vecn}\vdot{\vecn}{\vecy}\\
-  &= \vdot{\vecx}{\vecy}
-\end{split}
+  G_2(\vecv, \vecll, \vech(\vecv, \vecll)) = G_2(\vecll, \vecv, \vech(\vecll, \vecv)) \ .
 \end{equation}
-and therefore
-\begin{equation}\label{eq:khr}
-  \vdot{\vecx}{\vech(\vecx, \vecy)} =
-  \frac{1 + \vdot{\vecy}{\vecx}}{\lVert \vecx + \vecy \rVert} =
-  \frac{\vdot{\vecyy}{\vecyy} + \vdot{\vecyy}{\vecxx}}{\lVert \vecyy + \vecxx \rVert} =
-  \vdot{\vecyy}{\vech(\vecyy,\vecxx)}
+from which Equation~\eqref{eq:shadowing_symmetry} follows directly.
+
+\paragraph*{Fresnel term.}
+
+The Fresnel term $F(\vecv, \vech)$ in the microfacet model accounts for the proportion of light reflected vs. refracted at the micro-surface. We have to verify that
+\begin{equation}
+F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecll, \vech(\vecll, \vecv)) \ .
 \end{equation}
-as well as
-\begin{equation}\label{eq:kht}
-  \vdot{\vecx}{\vecht(\vecx, \vecy)} =
-  \frac{-\etax  -\etay \vdot{\vecx}{\vecy}}{\lVert -\etax\vecx  -\etay\vecy \rVert} =
-  \frac{-\etax \vdot{\vecyy}{\vecyy} - \etay\vdot{\vecxx}{\vecyy}}{\lVert -\etax\vecxx  -\etay\vecyy \rVert} =
-  \vdot{\vecyy}{\vech(\vecyy,\vecxx)}.
+Reflection symmetry of the Fresnel term implies
+\begin{equation}
+  F(\vecvv, \vech(\vecvv, \vecl)) = F(\vecv, \vech(\vecv, \vecll))
 \end{equation}
-As the Fresnel term takes the form $F(\vecx, \vecy) = F\left(\vdot{\vecx}{\vecy}\right)$, we can conclude
+and reciprocity symmetry further implies
 \begin{equation}
-    F\left(\vecvv, \vech(\vecvv, \vecl)\right) = F\left(\vdot{\vecvv}{\vech(\vecvv, \vecl)}\right) =
-    F\left(\vdot{\vecll}{\vech(\vecll,\vecv)}\right) = F\left(\vecll, \vech(\vecll,\vecv)\right).
+  F(\vecv, \vech(\vecv, \vecll)) = F(\vecll, \vech(\vecll, \vecv)) \ .
 \end{equation}
-Getting back to the geometry term, we first note that we can rewrite $G_1(\vecv, \vech)$ as $G_1(\vecv, \vech(\vecv, \vecl))$.
-For ease, we change the notation to $G_1(\vecv, \vecl)$ instead of $G_1(\vecv, \vech)$, and $G_1(\vecl, \vecv)$ instead of $G_1(\vecl, \vech)$.
-It follows that using Equations \eqref{eq:khr} and \eqref{eq:kht} along with Equations \eqref{eq:nh} and \eqref{eq:nk} und using $\vech(\vecx,\vecy) = \vech(\vecy,\vecx)$ we get
-\begin{equation}
-  \begin{split}
-  G_1(\vecvv, \vecl) &=
-  G_{1, \text{vc}}\left( \vdot{\vecvv}{\vecn}, \vdot{\vecn}{\vech(\vecvv, \vecl)}, \vdot{\vecvv}{\vech(\vecvv, \vecl)} \right)\\
-  &=  G_{1, \text{vc}}\left( \vdot{\vecv}{\vecn}, \vdot{\vecn}{\vech(\vecll, \vecv)}, \vdot{\vecll}{\vech(\vecll, \vecv)} \right) \\
-  &=  G_{1, \text{vc}}\left( \vdot{\vecv}{\vecn}, \vdot{\vecn}{\vech(\vecv, \vecll)}, \vdot{\vecll}{\vech(\vecv, \vecll)} \right) \\
-  &= G_1(\vecv, \vecll)
-  \end{split}
-\end{equation}
-and similarly $G_1(\vecl, \vecvv) = G_1(\vecll, \vecv)$ for the V-cavity masking function.
+from which the desired equality follows.
+
+\paragraph*{Summary.}
+
 In summary, we can conclude the reciprocity symmetry of the BRDF
@@
   f_r(\vecvv, \vecl) &=
-  \frac{D(\vech(\vecvv, \vecl)) \, G_{2}(\vecvv, \vecl) \, F\left(\vecvv, \vech(\vecvv, \vecl)\right)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn} }\\
+  \frac{D(\vech(\vecvv, \vecl)) \, G_{2}(\vecvv, \vecl) \, F\left(\vecvv, \vech(\vecvv, \vecl)\right)}{4 \avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} }\\
   &=
-  \frac{D(\vech(\vecll, \vecv)) \, G_{2}(\vecll, \vecv) \, F\left(\vecll, \vech(\vecll,\vecv)\right)}{4 \vdot{\vecv}{\vecn} \vdot{\vecll}{\vecn} }\\
+  \frac{D(\vech(\vecll, \vecv)) \, G_{2}(\vecll, \vecv) \, F\left(\vecll, \vech(\vecll,\vecv)\right)}{4 \avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} }\\
   &=
@@
 \end{equation}
+
+\section{Energy Conservation of MRM}
+
+\label{sec:energy_conservation}
+
+We now verify that the retro-reflective microfacet model fulfills energy conservation, i.e. that the directional albedos fulfill
+\begin{equation}
+  \rho_{r, \text{retro}}(\vecv) + \rho_{t, \text{retro}}(\vecv) \leq 1.
+\end{equation}
+
 For the directional albedos we simply have
@@
   \rho_{r, \text{retro}}(\vecv) =
-  \int_{\Omega_{\vecv}} f_{r, \text{retro}}(\vecv, \vecl) \vdot{\vecl}{\vecn} d\vecl =
-  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \vdot{\vecl}{\vecn} d\vecl = \rho_r(\vecvv),\\
+  \int_{\Omega_{\vecv}} f_{r, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
+  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_r(\vecvv),\\
   \rho_{t, \text{retro}}(\vecv) =
-  \int_{\Omega_{-\vecv}} f_{t, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} d\vecl =
-  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} d\vecl = \rho_t(\vecvv),
+  \int_{\Omega_{-\vecv}} f_{t, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
+  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_t(\vecvv),
   \end{split}
@@
 %% \end{equation}
+
 We further note that under our requirement of a reflection-symmetric microfacet normal distribution, the geometric
@@
 
+\newpage
+
 \section{Results and Discussion}
@@
 
-Figure~\ref{fig:data_fit} shows a fit of the retro-reflective MRM model (in red) to the measured BRDF of retro-reflective tape (in blue) from \cite{BelcourBackvector}.
-This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
-The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel with $F_0=0.19$ and $F_{82} = 0$.
-
-\begin{figure}[!b]
+\begin{figure}[!t]
   \begin{center}
@@
 
-\newpage
-
-\begin{figure}[!!t]
+\begin{figure}[!b]
   \centering
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_0_cropped.png}
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_45_cropped.png}
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/classic_angle_90_cropped.png} \\
-    \vspace{0.06cm}
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_0_cropped.png}
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_45_cropped.png}
-    \includegraphics[width=0.3\linewidth]{figures/conductor/cropped/retro_angle_90_cropped.png}
-  \captionsetup{font=small}
-  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction.
-The relative angle between the view and light direction is $0^\circ$ (left), $45^\circ$ (middle), and $90^\circ$ (right).
-\vspace{-0.5cm}}
+  \begin{tikzpicture}[font=\scriptsize ]
+    \node[inner sep=1pt] (GGX0) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/classic_angle_0_cropped.png}};
+    \node[inner sep=1pt,right = 0cm of GGX0] (GGX45) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/classic_angle_45_cropped.png}};
+    \node[inner sep=1pt,right = 0cm of GGX45] (GGX90) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/classic_angle_90_cropped.png}};
+    \node[inner sep=1pt,below = 0cm of GGX0] (MRM0) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/retro_angle_0_cropped.png}};
+    \node[inner sep=1pt,right = 0cm of MRM0] (MRM45) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/retro_angle_45_cropped.png}};
+    \node[inner sep=1pt,right = 0cm of MRM45] (MRM90) {\includegraphics[width=0.3\linewidth,frame]{figures/conductor/cropped/retro_angle_90_cropped.png}};
+    \node[inner sep=1pt,above = 2pt of GGX0] { \textbf{$\Delta\theta_{vl}=0^\circ$} };
+    \node[inner sep=1pt,above = 2pt of GGX45] { \textbf{$\Delta\theta_{vl}=45^\circ$} };
+    \node[inner sep=1pt,above = 2pt of GGX90] { \textbf{$\Delta\theta_{vl}=90^\circ$} };
+    \node[inner sep=1pt,left = 1pt of GGX0, anchor=south,rotate=90] { GGX };
+    \node[inner sep=1pt,left = 1pt of MRM0, anchor=south,rotate=90] { MRM };
+  \end{tikzpicture}
+  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction, with relative angle between the view and light direction $\Delta\theta_{vl} \in \{0^\circ, 45^\circ, 90^\circ\}$.}
   \label{fig:conductor_renders}
@@
 
-\begin{figure}[!b]
+\begin{figure}[!htb]
   \centering
-  \begin{tikzpicture}[font=\footnotesize]
+  \begin{tikzpicture}[font=\scriptsize ]
     \node[inner sep=1pt] (Class1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }};
-    \node[inner sep=1pt,right = 0cm of Class1] (MRM1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }};
-    \node[inner sep=1pt,right = 0cm of MRM1] (Mixed1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png }};
-
-    \node[inner sep=1pt,below = 0cm of Class1] (Class15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }};
-    \node[inner sep=1pt,right = 0cm of Class15] (MRM15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png }};
-    \node[inner sep=1pt,right = 0cm of MRM15] (Mixed15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }};
-
+    \node[inner sep=1pt,right = 0cm of Class1] (MRM1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1_cropped.png }};
+    \node[inner sep=1pt,right = 0cm of MRM1] (Mixed1) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1_cropped.png }};
+    \node[inner sep=1pt,below = 0cm of Class1] (Class15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/classic_ior_1.5_cropped.png }};
+    \node[inner sep=1pt,right = 0cm of Class15] (MRM15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroRT_ior_1.5_cropped.png }};
+    \node[inner sep=1pt,right = 0cm of MRM15] (Mixed15) {\includegraphics[width=0.3\linewidth,frame]{figures/dielectric/cropped/retroR_ior_1.5_cropped.png }};
     \node[inner sep=1pt,above = 2pt of Class1] { \textbf{Classical Microfacet} };
@@
     \node[inner sep=1pt,above = 0.5pt of Mixed1] { \textbf{and Classical (BTDF)} };
-
     \node[inner sep=1pt,left = 1pt of Class1, anchor=south,rotate=90] { $\mathbf{\boldsymbol{\eta} = 1}$ };
@@
   \end{tikzpicture}
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1_cropped_label.png }
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1_cropped_label.png }
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1_cropped_label.png } \\
-    % \vspace{0.06cm}
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/classic_ior_1.5_cropped_label.png }
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroRT_ior_1.5_cropped_label.png}
-    % \includegraphics[width=0.3\linewidth]{figures/dielectric/cropped/retroR_ior_1.5_cropped_label.png }
   \caption{Renders of a glass shaderball illuminated from the side.
@@
 
-\newpage
+Figure~\ref{fig:data_fit} shows a fit of the retro-reflective MRM model (in red) to the measured BRDF of retro-reflective tape (in blue) from \cite{BelcourBackvector}.
+This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
+The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel with $F_0=0.19$ and $F_{82} = 0$.
 
@@
 
-\paragraph{Further work.}
-For future work it would be interesting to further validate the MRM retro-reflective microfacet model against measured retro-reflective materials.
-
 \bibliographystyle{jcgt}

### 2025-12-22 article.tex
Cleanup, add references.

@@
 For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some examples).
-Such materials are typically designed to be retroreflective via a substructure of elements which preferentially scatter light backwards.
-A number of CG models have been proposed to model retro-reflective materials, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.
+Realistic rendering of such materials is important for design and visualization purposes, and in a predictive rendering context for example for simulation of visibility in training of autonomous vehicles or for safety certification of road markings.
 
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
-We present here a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets all these requirements.
+Materials are typically designed to be retroreflective via a manufactured substructure of elements (for example sheets of micro-prisms, or glass beads) which preferentially scatter light backwards \cite{Burgess2011}.
+A number of CG models have been proposed to model such materials, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.
+
+\paragraph{Previous works.}
+Modeling of retro-reflectivity in the surface scattering function has received relatively little attention so far.
+For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets.
+However, their model only provides pure retro-reflection which would lack artistic control.
+They later extended this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}.
+However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
+\citet{Hericz2016} proposed a retro-reflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.
+
+\paragraph{Our model.}
+
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving and reciprocal, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
+We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements.
 It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
 
-\paragraph{Previous works.}
-Modeling of retro-reflectivity in the surface scattering function has received little attention so far.
-For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets.
-However, their model only provides pure retro-reflection which would lack artistic control.
-They later extended this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\citet{Guo2018Retro}.
-However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
-
-\paragraph{Our model.} In this document, we present a model based on the previously published ``back-vector'' formulation~\cite{BelcourBackvector} which meets production requirements.
-It is also a simple modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
-
-We detail the correct usage and derivation of this retro-reflective BSDF model within the microfacet framework~\cite{WalterMicrofacet,HeitzMicrofacet}.
-We first recall some basic principles from microfacet theory (Section~\ref{sec:microfacets}), then show how this model can be steered from forward- to retro-reflection (Section~\ref{sec:backvector}), and further provide proof that the resulting BSDF is fully plausible (Section~\ref{sec:plausibility}).
+In Section \ref{sec:back_vector_microfacets}, we detail the derivation of MRM as a modification of the standard microfacet model~\cite{WalterMicrofacet,HeitzMicrofacet} using the back-vector as microfacet normal selector.
+To check the plausibility of the model, in Section \ref{sec:reciprocity} we verify reciprocity symmetry and in Section \ref{sec:energy_conservation} we verify energy conservation.
 In Section~\ref{sec:implementation}, we sketch how the model can be implemented in practice, which is done trivially by reusing existing microfacet BSDF implementations.
-Finally, we provide some visual results and discussion of the behavior (Section~\ref{sec:results}).
+Finally in Section~\ref{sec:results}, we provide a comparison with measured data and some visual results.
 
 \section{Back-vector Modification of Microfacet Models}
-\label{sec:microfacets}
+
+\label{sec:back_vector_microfacets}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[!tb]
   \begin{center}
-    \includegraphics[width=0.3\textwidth]{figures/backvector.pdf}
-    \includegraphics[width=0.3\textwidth]{figures/backvector_refract.pdf}
+    \includegraphics[width=0.4\textwidth]{figures/backvector.pdf}
+    \includegraphics[width=0.4\textwidth]{figures/backvector_refract.pdf}
   \end{center}
-  \caption{Geometry of the back-vector for reflection and refraction, illustrated in two dimensions.
+  \caption{Geometry of the back-vector for reflection and refraction in MRM, illustrated in two dimensions.
 The view vector $\vecv$ is reflected in a ``virtual mirror'' to obtain $\vecvv$, which is then used to compute the half-vector with the light vector $\vecl$.
@@
 
-As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a virtual plane aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$.
-Intuitively, this is equivalent to observing the microfacet reflection through a ``virtual mirror'', thus turning glossy forward-reflection into glossy retro-reflection.
-While there is no immediate physical foundation to this construction, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
+As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual mirror'' aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$, thus turning glossy forward-reflection into glossy retro-reflection.
+\citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface.
+Though here, there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
+
 We now detail the mathematical consequences of using the back-vector as microfacet selector.
@@
 
+\enlargethispage{1\baselineskip}
+
 The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry.
-It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction.
-In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
+It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction: \footnote{This is the seperable form commonly used in microfacet models; see Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
+non-separable shadowing-masking functions.
+In general, $G_{2}(\vecv, \vecl, \vech)$ is some function of $G_1(\vecv, \vech)$ and $G_1(\vecl, \vech)$, and thus reciprocity still holds.}
+%In the case of independent shadowing and masking, it is common to use the separable form\footnote{See Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
 non-separable shadowing-masking functions.
@@
 \end{equation}
-Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
-$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
-\begin{equation}\label{eq:visible}
-  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
-\end{equation}
-Popular choices for $G_1$ with this property are the Smith masking function
+%Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
+%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
+%\begin{equation}\label{eq:visible}
+%  \avdot{\vecv}{\vecn} = \int_{\Omega_{\vecv}} G_1(\vecv, \vech) \cldot{\vecv}{\vech} D(\vech) \diff\vech.
+%\end{equation}
+Popular choices for $G_1$
+%with this property
+are the Smith masking function
 \begin{equation} \label{eq:smith_masking}
@@
 
-We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution.
+We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution, and thus replace  $G_{2}(\vecv, \vecl, \vech)$  with $G_{2}(\vecvv, \vecl, \vech)$ in the retro-reflective microfacet model.
 With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
-For V-cavity masking, we take $\vecvv$ as view direction and, consequently, $\vecb$ as microfacet normal.
+For V-cavity masking, taking $\vecvv$ as view direction simply corresponds to selecting $\vecb$ as microfacet normal.
 %\todo{\tiny Laurent: I think we could extract the shadowing from Matthias derivation here.
@@
 \paragraph*{Final form.}
-It follows from all this that the final retro-reflective microfacet model can be written simply as
+It follows from the above back-vector modifications that our retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
 \begin{equation} \label{eq:retro_microfacet_model}
-  %  f_\text{retro}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
-  f_\text{retro}(\vecv, \vecl) =
+  %  f_\text{MRM}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
+  f_\text{MRM}(\vecv, \vecl) =
   \begin{cases}
-    f_{r, \text{retro}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
-    f_{t, \text{retro}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
+    f_{r, \text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
+    f_{t, \text{MRM}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
   \end{cases}
 \end{equation}
-In the following section we show in detail that this model is plausible, i.e. that symmetry and energy conservation requirements are fulfilled.
-Note that for this we impose one technical assumption on $D$:
-we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
-
 \begin{figure}[!tb]
   \centering
-  \includegraphics[width=0.83\linewidth]{figures/brdf_plots.pdf}
+  \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
   \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
@@
 
-Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue).
+To gain some quantitative insight into the behavior of the MRM model, Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue).
 For both MRM and GGX, the Fresnel factor is set to 1.
+\enlargethispage{1\baselineskip}
 The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
@@
 
-We briefly show here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills symmetry and energy conservation requirements.
+\section{Plausibility of MRM}
 
-\section{Reciprocity of MRM}
+We briefly show here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e.
+that it fulfills reciprocity symmetry and energy conservation requirements.
+
+\enlargethispage{0.5\baselineskip}
+
+\subsection{Reciprocity}
 
@@
 
-We first verify that the retro-reflective microfacet model fulfills reciprocity symmetry, i.e.
+We first verify that the MRM microfacet model fulfills reciprocity symmetry, i.e. that:
 \begin{equation}
-  f_{r,\text{retro}}(\vecv, \vecl) = f_{r,\text{retro}}(\vecl, \vecv) \ .
+  f_{\text{MRM}}(\vecv, \vecl) = f_{\text{MRM}}(\vecl, \vecv) \ , \quad f_{t,\text{MRM}}(\vecv, \vecl) = \frac{\etav^2}{\etal^2}f_{t,\text{MRM}}(\vecl, \vecv) \ .
 \end{equation}
@@
 We will verify explicitly that the retro-reflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
+Note that for this we will need to impose one technical assumption on $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
 
@@
 as can be verified with Equation~\eqref{eq:vc_masking}.
-In the case of Smith masking, this also holds provided that $D$ is reflection-symmetric around the normal (which we assumed), as follows from Equation~(\eqref{eq:smith_masking}.
+In the case of Smith masking, this also holds provided that $D$ is reflection-symmetric around the normal (which we assumed), as follows from Equation~\eqref{eq:smith_masking}.
 Then it follows from standard reciprocity that
@@
 \begin{equation}
-  \begin{split}
-  f_{r,\text{retro}}(\vecv, \vecl) =
-  f_r(\vecvv, \vecl) &=
-  \frac{D(\vech(\vecvv, \vecl)) \, G_{2}(\vecvv, \vecl) \, F\left(\vecvv, \vech(\vecvv, \vecl)\right)}{4 \avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} }\\
-  &=
-  \frac{D(\vech(\vecll, \vecv)) \, G_{2}(\vecll, \vecv) \, F\left(\vecll, \vech(\vecll,\vecv)\right)}{4 \avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} }\\
-  &=
-  f_r(\vecll, \vecv) =
-  f_{r,\text{retro}}(\vecl, \vecv).
-  \end{split}
+  f_{r,\text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl) =f_r(\vecll, \vecv) =  f_{r,\text{MRM}}(\vecl, \vecv).
 \end{equation}
+
+%\begin{equation}
+%  \begin{split}
+%  f_{r,\text{MRM}}(\vecv, \vecl) =
+%  f_r(\vecvv, \vecl) &=
+%  \frac{D(\vech(\vecvv, \vecl)) \, G_{2}(\vecvv, \vecl) \, F\left(\vecvv, \vech(\vecvv, \vecl)\right)}{4 \avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} }\\
+%  &=
+%  \frac{D(\vech(\vecll, \vecv)) \, G_{2}(\vecll, \vecv) \, F\left(\vecll, \vech(\vecll,\vecv)\right)}{4 \avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} }\\
+%  &=
+%  f_r(\vecll, \vecv) =
+%  f_{r,\text{MRM}}(\vecl, \vecv).
+%  \end{split}
+%\end{equation}
 Analogously, we obtain the reciprocity symmetry requirement of the BTDF
 \begin{equation}
-  \begin{split}
-    &f_{t,\text{retro}}(\vecv, \vecl)
-    =  f_t(\vecvv, \vecl)\\
-    &= \frac{\avdot{\vecvv}{\vecht(\vecvv, \vecl)}\avdot{\vecl}{\vecht(\vecvv, \vecl)}\etav^2 D(\vecht(\vecvv, \vecl)) \, G_{2}(\vecvv,\vecl) \, \left(1 - F(\vecvv, \vecht(\vecvv, \vecl)) \right)}
-    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecvv}{\vecht(\vecvv, \vecl)} + \etal \vdot{\vecl}{\vecht(\vecvv, \vecl)}\right)^2} \\
-    &= \frac{\avdot{\vecv}{\vecht(\vecv, \vecll)}\avdot{\vecll}{\vecht(\vecv, \vecll)}\etav^2 D(\vecht(\vecll, \vecv)) \, G_{2}(\vecll,\vecv) \, \left(1 - F(\vecll, \vecht(\vecll, \vecv)) \right)}
-    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \vdot{\vecv}{\vecht(\vecv, \vecll)} + \etal \vdot{\vecll}{\vecht(\vecv, \vecll)}\right)^2} \\
-    &=
-    \frac{\etav^2}{\etal^2}f_t(\vecll, \vecv) =
-    \frac{\etav^2}{\etal^2}f_{t,\text{retro}}(\vecl, \vecv).
-  \end{split}
+    f_{t,\text{MRM}}(\vecv, \vecl) =  f_t(\vecvv, \vecl) = \frac{\etav^2}{\etal^2}f_t(\vecll, \vecv) = \frac{\etav^2}{\etal^2}f_{t,\text{MRM}}(\vecl, \vecv).
 \end{equation}
 
-\section{Energy Conservation of MRM}
+%\begin{equation}
+%  \begin{split}
+%    &f_{t,\text{MRM}}(\vecv, \vecl)
+%    =  f_t(\vecvv, \vecl)\\
+%    &= \frac{\avdot{\vecvv}{\vecht(\vecvv, \vecl)}\avdot{\vecl}{\vecht(\vecvv, \vecl)}\etav^2 D(\vecht(\vecvv, \vecl)) \, G_{2}(\vecvv,\vecl) \, \left(1 - F(\vecvv, %\vecht(\vecvv, \vecl)) \right)}
+%    {\avdot{\vecvv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecvv}{\vecht(\vecvv, \vecl)} + \etal \vdot{\vecl}{\vecht(\vecvv, \vecl)}\right)^2} \\
+%    &= \frac{\avdot{\vecv}{\vecht(\vecv, \vecll)}\avdot{\vecll}{\vecht(\vecv, \vecll)}\etav^2 D(\vecht(\vecll, \vecv)) \, G_{2}(\vecll,\vecv) \, \left(1 - F(\vecll, %\vecht(\vecll, \vecv)) \right)}
+%    {\avdot{\vecv}{\vecn} \avdot{\vecll}{\vecn} \left(\etav \vdot{\vecv}{\vecht(\vecv, \vecll)} + \etal \vdot{\vecll}{\vecht(\vecv, \vecll)}\right)^2} \\
+%    &=
+%    \frac{\etav^2}{\etal^2}f_t(\vecll, \vecv) =
+%    \frac{\etav^2}{\etal^2}f_{t,\text{MRM}}(\vecl, \vecv).
+%  \end{split}
+%\end{equation}
+
+\subsection{Energy Conservation}
 
@@
 \begin{equation}
-  \rho_{r, \text{retro}}(\vecv) + \rho_{t, \text{retro}}(\vecv) \leq 1.
+  \rho_{r, \text{MRM}}(\vecv) + \rho_{t, \text{MRM}}(\vecv) \leq 1.
 \end{equation}
-
 For the directional albedos we simply have
@@
   \begin{split}
-  \rho_{r, \text{retro}}(\vecv) =
-  \int_{\Omega_{\vecv}} f_{r, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
+  \rho_{r, \text{MRM}}(\vecv) =
+  \int_{\Omega_{\vecv}} f_{r, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
   \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_r(\vecvv),\\
-  \rho_{t, \text{retro}}(\vecv) =
-  \int_{\Omega_{-\vecv}} f_{t, \text{retro}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
-  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_t(\vecvv),
+  \rho_{t, \text{MRM}}(\vecv) =
+  \int_{\Omega_{-\vecv}} f_{t, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
+  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_t(\vecvv) \ .
   \end{split}
 \end{equation}
-i.e. we get the albedo of the regular microfacet BSDF for the reflected direction.
+That is, the directional albedo of MRM is equivalent to the directional albedo of the regular microfacet BSDF, but evaluated at the reflected direction.
 As the original microfacet BSDF fulfills energy conversion, the retro-reflective modification does, too.
@@
 %% \end{equation}
-
 We further note that under our requirement of a reflection-symmetric microfacet normal distribution, the geometric
@@
 Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retro-reflection are the same, i.e.
-$\rho_{r, \text{retro}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{retro}}(\vecv) = \rho_t(\vecv)$.
+$\rho_{r, \text{MRM}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{MRM}}(\vecv) = \rho_t(\vecv)$.
 
@@
 \item
-  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for energy loss in the sense of \cite{KelemenBRDF} can be realized using the same data tables.
+  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Hill2017PBR} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables.
+The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.
+\item An implementation may optionally wish to provide a blend between regular and retro-reflective microfacet behavior, to model materials with partial retro-reflective properties.
+This can be achieved simply by blending $f_{r, \text{MRM}}$ and $f_r$, or alternatively via a stochastic selection (e.g.
+one-sample MIS \cite{Veach1998}) between the original $\vecv$ and reflected $\vecvv$ view directions during sampling.
 \end{itemize}
@@
 
+\enlargethispage{0.5\baselineskip}
+
 \begin{figure}[!t]
@@
   \end{center}
+  \vspace{-0.5cm}
   \caption{Measured BRDF of retro-reflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
@@
 This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
-The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel with $F_0=0.19$ and $F_{82} = 0$.
+The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel \cite{OpenPBR2024} with $F_0=0.19$ and $F_{82} = 0$.
 
@@
 The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
-When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{retro}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{MRM}}$ yields more intuitive results, in particular for an index of refraction close to 1.
 
 Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
-Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{retro}}$ plus retro-reflective BTDF $f_{t,\text{retro}}$ (middle), and retro-reflective BRDF $f_{r,\text{retro}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
+Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{MRM}}$ plus retro-reflective BTDF $f_{t,\text{MRM}}$ (middle), and retro-reflective BRDF $f_{r,\text{MRM}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
 
@@
 We have presented a retro-reflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
+We have shown that the model fulfills symmetry and energy conservation requirements, making it a plausible microfacet model.
+This model can be used to represent materials with strong retro-reflective properties, such as certain fabrics or retro-reflective coatings,  with a reasonable degree of accuracy and very efficiently.
 The model is straightforward to implement in existing renderers supporting microfacet BSDFs, as it only requires reflecting the view direction about the surface normal before evaluating the standard microfacet model.
-We have shown that the model fulfills symmetry and energy conservation requirements, making it a plausible microfacet model.
-The retro-reflective microfacet model can be used to represent materials with strong retro-reflective properties, such as certain fabrics or retro-reflective coatings.
+
+Due to its simplicity and ease of integration into existing microfacet-based rendering systems, the MRM model has been proposed as a standard component in the OpenPBR material model \cite{OpenPBR2024} (and correspondingly in the MaterialX modeling framework \cite{Smythe2016}).
 

### 2025-12-22 code/brdf_plotting_with_data.py

@@ -111,9 +111,9 @@ for idx, (tL, data_points) in enumerate(sorted(brdf_dict_simple.items())):
     # Plot BRDF values vs theta_V (in degrees)
     axes[idx].clear()  # Clear previous plot
     axes[idx].plot(theta_V_vals_deg, brdf_vals, 'o-', linewidth=1.5, markersize=4)
-    axes[idx].set_xlabel('$\\theta_V$ (degrees)', fontsize=12)
-    axes[idx].set_ylabel('BRDF value', fontsize=12)
-    axes[idx].set_title(f'$\\theta_L = {tL_deg:.2f}^\\circ$', fontsize=14)
+    axes[idx].set_xlabel('$\\theta_V$ (degrees)', fontsize=16)
+    axes[idx].set_ylabel('BRDF value', fontsize=16)
+    axes[idx].set_title(f'$\\theta_L = {tL_deg:.2f}^\\circ$', fontsize=16)
     axes[idx].set_xlim(-90, 90)
     axes[idx].set_ylim(0.001, 50)
     axes[idx].set_yscale('log')

### 2025-12-22 article.tex
Improve text.

@@
 
-\section{Back-vector Modification of Microfacet Models}
+\section{MRM as Back-vector Modification of Microfacet Models}
 
@@
 
-To achieve retro-reflection, we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
+To achieve retro-reflection, in MRM we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
 \begin{equation}
@@
 \paragraph*{Final form.}
-It follows from the above back-vector modifications that our retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
+It follows from the above back-vector modifications that our MRM retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
 \begin{equation} \label{eq:retro_microfacet_model}
@@
 The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.
-\item An implementation may optionally wish to provide a blend between regular and retro-reflective microfacet behavior, to model materials with partial retro-reflective properties.
+\item An implementation may optionally wish to provide a blend between regular and retro-reflective microfacet behavior, to model materials with partial retro-reflective properties (for example this is required in OpenPBR \cite{OpenPBR2024}).
 This can be achieved simply by blending $f_{r, \text{MRM}}$ and $f_r$, or alternatively via a stochastic selection (e.g.
@@
 }
+
 Eval_result eval_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
@@
 When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{MRM}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+An implementation could reasonably choose either this formulation or the full retro-reflective BTDF, depending on the desired effect.
 

### 2025-12-22 article.tex
Caption tweak.

@@
     \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
-  \caption{Comparison of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
+  \caption{Renders of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
 The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak in the MRM model.}
@@
 Rendering of retro-reflective materials is useful in a variety of contexts.
-For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some examples).
+For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some example renders).
 Realistic rendering of such materials is important for design and visualization purposes, and in a predictive rendering context for example for simulation of visibility in training of autonomous vehicles or for safety certification of road markings.

### 2025-12-22 article.tex
Direct Kulla & Conty reference.

@@
 \item
-  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Hill2017PBR} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables.
+  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Kulla17} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables.
 The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.

### 2025-12-23 article.tex
Fix hyphenation

@@
   \caption{Renders of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
-The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak in the MRM model.}
+The lighting in all cases is aligned with the view direction, exhibiting a strong retroreflective peak in the MRM model.}
    \vspace{-0.5cm}
@@
 
-Rendering of retro-reflective materials is useful in a variety of contexts.
+Rendering of retroreflective materials is useful in a variety of contexts.
 For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some example renders).
@@
 \paragraph{Previous works.}
-Modeling of retro-reflectivity in the surface scattering function has received relatively little attention so far.
-For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets.
-However, their model only provides pure retro-reflection which would lack artistic control.
-They later extended this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}.
-However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
-\citet{Hericz2016} proposed a retro-reflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.
+Modeling of retroreflectivity in the surface scattering function has received relatively little attention so far.
+For example, \citet{Guo2017Retro} proposed a model to incorporate the retroreflectivity of prismatic sheets.
+However, their model only provides pure retroreflection which would lack artistic control.
+They later extended this model to account for imperfect retroreflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}.
+However, due to the coupling of retroreflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
+\citet{Hericz2016} proposed a retroreflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.
 
@@
 We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements.
-It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.
+It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible ive result.
 
@@
 In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
-In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retro-reflection.
+In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retroreflection.
 %Before detailing the impact of such a change, we review some standard properties of microfacet models.
@@
 
-To achieve retro-reflection, in MRM we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
+To achieve retroreflection, in MRM we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
 \begin{equation}
@@
 
-As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual mirror'' aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$, thus turning glossy forward-reflection into glossy retro-reflection.
+As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual mirror'' aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$, thus turning glossy forward-reflection into glossy retroreflection.
 \citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface.
@@
 \paragraph*{Microfacet reflection term.}
-There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
+There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retroreflective material measurements for a reflection back-vector based BRDF.
 As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead.
@@
 
-We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution, and thus replace  $G_{2}(\vecv, \vecl, \vech)$  with $G_{2}(\vecvv, \vecl, \vech)$ in the retro-reflective microfacet model.
+We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution, and thus replace  $G_{2}(\vecv, \vecl, \vech)$  with $G_{2}(\vecvv, \vecl, \vech)$ in the retroreflective microfacet model.
 With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration.
@@
   \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
-  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
+  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
 The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$.
@@
 In this case.
-both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retro-reflection direction respectively for GGX and MRM).
+both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retroreflection direction respectively for GGX and MRM).
 
@@
 
-We briefly show here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e.
+We briefly show here that the retroreflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e.
 that it fulfills reciprocity symmetry and energy conservation requirements.
@@
 For brevity, in analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
-We will verify explicitly that the retro-reflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
+We will verify explicitly that the retroreflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
 Note that for this we will need to impose one technical assumption on $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
@@
 
-We now verify that the retro-reflective microfacet model fulfills energy conservation, i.e. that the directional albedos fulfill
+We now verify that the retroreflective microfacet model fulfills energy conservation, i.e. that the directional albedos fulfill
 \begin{equation}
@@
 That is, the directional albedo of MRM is equivalent to the directional albedo of the regular microfacet BSDF, but evaluated at the reflected direction.
-As the original microfacet BSDF fulfills energy conversion, the retro-reflective modification does, too.
+As the original microfacet BSDF fulfills energy conversion, the retroreflective modification does, too.
 %% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
@@
 configurations for $\vecv$ and $\vecvv$ are identical.
-Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retro-reflection are the same, i.e.
+Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retroreflection are the same, i.e.
 $\rho_{r, \text{MRM}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{MRM}}(\vecv) = \rho_t(\vecv)$.
@@
 
-Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):
+Given an implementation of a regular microfacet BSDF, extending it to retroreflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):
 
@@
 \item
-  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Kulla17} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables.
+  As the albedos of standard BSDF and retroreflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Kulla17} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables.
 The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.
-\item An implementation may optionally wish to provide a blend between regular and retro-reflective microfacet behavior, to model materials with partial retro-reflective properties (for example this is required in OpenPBR \cite{OpenPBR2024}).
+\item An implementation may optionally wish to provide a blend between regular and retroreflective microfacet behavior, to model materials with partial retroreflective properties (for example this is required in OpenPBR \cite{OpenPBR2024}).
 This can be achieved simply by blending $f_{r, \text{MRM}}$ and $f_r$, or alternatively via a stochastic selection (e.g.
@@
     language=GLSL,
-    caption=Pseudo code to implement the retro-reflective microfacet model using an existing regular microfacet BSDF implementation,
+    caption=Pseudo code to implement the retroreflective microfacet model using an existing regular microfacet BSDF implementation,
     label=lst:implementation,
@@
   \vspace{-0.5cm}
-  \caption{Measured BRDF of retro-reflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
+  \caption{Measured BRDF of retroreflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
   \label{fig:data_fit}
@@
   \end{tikzpicture}
-  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retro-reflective MRM material (bottom row) applied, for various camera angles and a fixed light direction, with relative angle between the view and light direction $\Delta\theta_{vl} \in \{0^\circ, 45^\circ, 90^\circ\}$.}
+  \caption{Renders of a metallic shaderball with either a classic GGX (top row) or retroreflective MRM material (bottom row) applied, for various camera angles and a fixed light direction, with relative angle between the view and light direction $\Delta\theta_{vl} \in \{0^\circ, 45^\circ, 90^\circ\}$.}
   \label{fig:conductor_renders}
@@
     \node[inner sep=1pt,above = 2pt of Class1] { \textbf{Classical Microfacet} };
-    \node[inner sep=1pt,above = 2pt of MRM1] { \textbf{Full retro-reflective MRM} };
+    \node[inner sep=1pt,above = 2pt of MRM1] { \textbf{Full retroreflective MRM} };
     \node[inner sep=1pt,above = 11.5pt of Mixed1] { \textbf{Mixed MRM (BRDF)} };
@@
 The index of refraction is $1.0$ (top) or $1.5$ (bottom).
-From left to right: classic microfacet BSDF (left), retro-reflective BSDF (middle), and combined retro-reflective BRDF and classic BTDF (right).
+From left to right: classic microfacet BSDF (left), retroreflective BSDF (middle), and combined retroreflective BRDF and classic BTDF (right).
 }
@@
 
-Figure~\ref{fig:data_fit} shows a fit of the retro-reflective MRM model (in red) to the measured BRDF of retro-reflective tape (in blue) from \cite{BelcourBackvector}.
+Figure~\ref{fig:data_fit} shows a fit of the retroreflective MRM model (in red) to the measured BRDF of retroreflective tape (in blue) from \cite{BelcourBackvector}.
 This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
-The MRM model provides a reasonably good match to the retro-reflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel \cite{OpenPBR2024} with $F_0=0.19$ and $F_{82} = 0$.
+The MRM model provides a reasonably good match to the retroreflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel \cite{OpenPBR2024} with $F_0=0.19$ and $F_{82} = 0$.
 
@@
 
-Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retro-reflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.
+Figure~\ref{fig:conductor_renders} shows renders of a shaderball with either a classic GGX or retroreflective MRM material applied, here assuming \emph{conducting} (i.e. metallic) microfacets.
 
@@
 
-The albedo is identical for forward- and retro-reflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retro-reflective variants into a BSDF.
-When constructing an artist-controllable retro-reflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{MRM}}$ yields more intuitive results, in particular for an index of refraction close to 1.
-An implementation could reasonably choose either this formulation or the full retro-reflective BTDF, depending on the desired effect.
+The albedo is identical for forward- and retroreflective BRDF as well as BTDF, which allows to arbitrarily combine forward- and retroreflective variants into a BSDF.
+When constructing an artist-controllable retroreflective dielectric BSDF, we observed that using $f_t$ instead of $f_{t, \text{MRM}}$ yields more intuitive results, in particular for an index of refraction close to 1.
+An implementation could reasonably choose either this formulation or the full retroreflective BTDF, depending on the desired effect.
 
 Figure~\ref{fig:dielectric_renders} demonstrates this in renders of a dielectric shaderball, where the IOR of the dielectric material is either $1.0$ (top row) or $1.5$ (bottom row).
-Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retro-reflective BRDF $f_{r,\text{MRM}}$ plus retro-reflective BTDF $f_{t,\text{MRM}}$ (middle), and retro-reflective BRDF $f_{r,\text{MRM}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
+Left to right we show three combinations of BRDF and BTDF: the classic microfacet BSDF $f_r$ and BTDF $f_t$ (left), the retroreflective BRDF $f_{r,\text{MRM}}$ plus retroreflective BTDF $f_{t,\text{MRM}}$ (middle), and retroreflective BRDF $f_{r,\text{MRM}}$ combined with the classic (non-retroreflective) BTDF $f_t$ (right).
 
@@
 
-We have presented a retro-reflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
+We have presented a retroreflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
 We have shown that the model fulfills symmetry and energy conservation requirements, making it a plausible microfacet model.
-This model can be used to represent materials with strong retro-reflective properties, such as certain fabrics or retro-reflective coatings,  with a reasonable degree of accuracy and very efficiently.
+This model can be used to represent materials with strong retroreflective properties, such as certain fabrics or retroreflective coatings,  with a reasonable degree of accuracy and very efficiently.
 The model is straightforward to implement in existing renderers supporting microfacet BSDFs, as it only requires reflecting the view direction about the surface normal before evaluating the standard microfacet model.

### 2025-12-23 article.tex
Typo fixes

@@
 %Mathematically, the BSDF is plausible if $G_{2}$ is symmetric w.r.t. to exchanging $\vecv$ and $\vecl$, fulfills
-%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible micro-surfaces \cite{HeitzMicrofacet}
+%$G_{2}(\vecv, \vech) \leq G_1(\vecv, \vecl)$, and if $G_1$ further establishes the correct projected area of visible microsurfaces \cite{HeitzMicrofacet}
 %\begin{equation}\label{eq:visible}
@@
 We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
-Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
+Similarly, for the transmissive microsurface BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
 \begin{equation}
@@
 The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry.
-It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction: \footnote{This is the seperable form commonly used in microfacet models; see Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
+It combines the visibility masking function $G_1(\vecv, \vech)$ for the view direction and the visibility shadowing function $G_1(\vecl, \vech)$ for the light direction: \footnote{This is the separable form commonly used in microfacet models; see Section 6 of \citet{HeitzMicrofacet} for a discussion of separable vs.
 non-separable shadowing-masking functions.
@@
 
-The Fresnel term $F(\vecv, \vech)$ in the microfacet model accounts for the proportion of light reflected vs. refracted at the micro-surface. We have to verify that
+The Fresnel term $F(\vecv, \vech)$ in the microfacet model accounts for the proportion of light reflected vs. refracted at the microsurface. We have to verify that
 \begin{equation}

### 2025-12-23 article.tex
Further minor fixes and cleanup.

@@
   \small
-  We describe the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible, energy-conserving retroreflective result.
+  We describe the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible retroreflective result which is energy-conserving and reciprocal by construction.
   \end{abstract}
@@
 Materials are typically designed to be retroreflective via a manufactured substructure of elements (for example sheets of micro-prisms, or glass beads) which preferentially scatter light backwards \cite{Burgess2011}.
-A number of CG models have been proposed to model such materials, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.
 
@@
 Modeling of retroreflectivity in the surface scattering function has received relatively little attention so far.
-For example, \citet{Guo2017Retro} proposed a model to incorporate the retroreflectivity of prismatic sheets.
+ A number of CG models have been proposed, but typically they are either not energy-conserving, not easily integrated into existing rendering systems, or too complex to be practical for production use.
+In \citet{Neumann1999} a retroreflective model based on a modified Phong lobe was presented, but it is not energy-conserving.
+\citet{Edwards2006} proposed an empirical retroreflective model, but it is not reciprocal.
+Neither of these models were compared to physical measurements.
+\citet{Guo2017Retro} proposed a model to incorporate the retroreflectivity of prismatic sheets.
 However, their model only provides pure retroreflection which would lack artistic control.
@@
 However, due to the coupling of retroreflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
-\citet{Hericz2016} proposed a retroreflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.
+\citet{Hericz2016} proposed a retroreflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF.
 
@@
 
-For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving and reciprocal, c) efficient and easy to implement, ideally a small modification to well-understood surface models.
+For practical purposes in visual effects, we are interested in a model which is a) visually plausible and follows measured observations reasonably well, b) provably energy-conserving and reciprocal, c) efficient and easy to implement.
 We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements.
@@
 where $D$ is the distribution of micronormals (NDF), normalized to fulfill $\int_\Omega D(\vech) (\vdot{\vecn}{\vech}) \diff\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term.
-The half-vector $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
+The half-vector $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (i.e.
+only microfacets with normals aligned with the half-vector will reflect light).
 Similarly, using the half-vector of refraction $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$ (where $\etav$, $\etal$ are the IORs of the media $\vecv$, $\vecl$ point to) as selector yields the corresponding BTDF:
@@
 \citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface.
-Though here, there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
+Though here there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
 
@@
 The term $G_{2}(\vecv,\vecl,\vechr)$ in the microfacet model accounts for shadowing and masking of microfacets.
-
 \enlargethispage{1\baselineskip}
-
 The shadowing-masking function $G_{2}(\vecv, \vecl, \vech)$ accounts for the geometrical occlusion of the incident and outgoing rays by the local microsurface geometry.
@@
   \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
-  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$.
-The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$.
-The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
+  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (red) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in \{45^\circ, 80^\circ\}$.
+The dashed lines are the roughness $0.1$ case, and the solid lines roughness $0.666$.
+The horizontal green line at $1/\pi$ is the Lambert BRDF.}
   \label{fig:brdf_plots}
@@
 
+\paragraph*{BRDF comparison.}
 To gain some quantitative insight into the behavior of the MRM model, Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue).
@@
 It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
-The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane.
+The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane as the light direction.
 In this case.
@@
 
-We briefly show here that the retroreflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e.
+We briefly show here that the retroreflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is physically plausible, i.e.
 that it fulfills reciprocity symmetry and energy conservation requirements.
@@
 We will verify explicitly that the retroreflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
-Note that for this we will need to impose one technical assumption on $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
+Note that for this we will need to impose one technical assumption on the NDF $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).
 
@@
 \end{equation}
-
 %\begin{equation}
@@
 We have presented a retroreflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
-We have shown that the model fulfills symmetry and energy conservation requirements, making it a plausible microfacet model.
+We have shown that the model fulfills reciprocity and energy conservation requirements, making it a physically plausible microfacet model.
 This model can be used to represent materials with strong retroreflective properties, such as certain fabrics or retroreflective coatings,  with a reasonable degree of accuracy and very efficiently.
@@
 Due to its simplicity and ease of integration into existing microfacet-based rendering systems, the MRM model has been proposed as a standard component in the OpenPBR material model \cite{OpenPBR2024} (and correspondingly in the MaterialX modeling framework \cite{Smythe2016}).
+
+\small
 

### 2025-12-23 code/brdf_plotting.py

@@ -68,7 +68,13 @@ def plot_brdf_data():
         index = get_index(theta_i, phi)
 
         # Set custom x-axis ticks
-        custom_ticks = np.arange(-90, 91, 45)  # Creates ticks from -90 to 90 with a step of 22.5
+        if theta_i == '45':
+            custom_ticks = [-90, -45, 0, 45, 90]
+        elif theta_i == '80':
+            custom_ticks = [-80, -45, 0, 45, 80]
+        else:
+            custom_ticks = [-80, -45, 0, 45, 80]
+
         axes[index].set_xticks(custom_ticks)
 
         # Set custom x-axis tick labels with a degree symbol
@@ -77,7 +83,7 @@ def plot_brdf_data():
 
         # Vertical lines at specified theta_o values
         for theta_line in [-80, -45, 0, 45, 80]:
-            axes[index].axvline(x=theta_line, color='lightgray', linestyle='dotted', linewidth=0.7)
+            axes[index].axvline(x=theta_line, color='gray', linestyle='dotted', linewidth=1)
 
         # Setting y-ticks for every 0.05 and adding horizontal grid lines
         #axes[index].set_yticks(np.arange(0, 0.28, 0.05))
@@ -93,13 +99,13 @@ def plot_brdf_data():
             linestyle = 'solid'
             show_legend = True
 
-        axes[index].plot(data['theta_o'], data['Lambert'], label='Lambert' if show_legend else None, color='gray',   linewidth=1.0, linestyle=linestyle)
+        axes[index].plot(data['theta_o'], data['Lambert'], label='Lambert' if show_legend else None, color='green',   linewidth=1.0, linestyle=linestyle)
         axes[index].plot(data['theta_o'], data['EON'],     label='EON'     if show_legend else None, color='orange', linewidth=1.0, linestyle=linestyle)
         axes[index].plot(data['theta_o'], data['GGX'],     label='GGX'     if show_legend else None, color='blue',   linewidth=1.0, linestyle=linestyle)
-        axes[index].plot(data['theta_o'], data['MRM'],     label='MRM'     if show_legend else None, color='green',  linewidth=1.0, linestyle=linestyle)
[…]

### 2025-12-23 MRM.tex
filename change

--- 2025-12-23 article.tex
@@
   \vspace*{-0.5cm}
-    \includegraphics[width=0.3\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
-    \includegraphics[width=0.3\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
-    \includegraphics[width=0.3\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
+    \includegraphics[width=0.31\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
+    \includegraphics[width=0.31\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
+    \includegraphics[width=0.31\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
     \vspace{0.06cm}
-    \includegraphics[width=0.3\linewidth]{figures/cone/retro_cone_cropped_label.png}
-    \includegraphics[width=0.3\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
-    \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
-  \caption{Renders of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
+    \includegraphics[width=0.31\linewidth]{figures/cone/retro_cone_cropped_label.png}
+    \includegraphics[width=0.31\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
+    \includegraphics[width=0.31\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
+  \caption{Renders of our MRM retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row).
 The lighting in all cases is aligned with the view direction, exhibiting a strong retroreflective peak in the MRM model.}
@@
   \small
-  We describe the ``Minimal Retroreflective Microfacet'' (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a visually plausible retroreflective result which is energy-conserving and reciprocal by construction.
+  We describe the Minimal Retroreflective Microfacet (MRM) model, which provides a simple scheme for modifying a pre-existing microfacet BSDF implementation to produce a physically plausible retroreflective result.
+Based on the previously published back-vector formulation, MRM can be implemented with minimal changes to existing microfacet code, making it easy to integrate into existing rendering systems.
+We provide verification of reciprocity and energy conservation, implementation details, and a comparison to measured data.
   \end{abstract}
@@
 For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some example renders).
+Materials are typically designed to be retroreflective via a manufactured substructure of elements (for example sheets of micro-prisms, or glass beads) which preferentially scatter light backwards \cite{Burgess2011}.
 Realistic rendering of such materials is important for design and visualization purposes, and in a predictive rendering context for example for simulation of visibility in training of autonomous vehicles or for safety certification of road markings.
-
-Materials are typically designed to be retroreflective via a manufactured substructure of elements (for example sheets of micro-prisms, or glass beads) which preferentially scatter light backwards \cite{Burgess2011}.
 
@@
 However, their model only provides pure retroreflection which would lack artistic control.
-They later extended this model to account for imperfect retroreflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}.
-However, due to the coupling of retroreflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model.
+They later extended this model to account for imperfect retroreflection with the addition of a microfacet surface above the prismatic sheet~\cite{Guo2018Retro}.
+However %, due to the coupling of retroreflection and rough refraction,
+they were unable to derive a close form solution for it. %and relied on fitting of the ground truth model.
 \citet{Hericz2016} proposed a retroreflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF.
+
+\enlargethispage{0.1\baselineskip}
 
@@
 We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements.
-It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible ive result.
+It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible result.
 
@@
 
-We have presented a retroreflective microfacet model (MRM) based on a simple modification of the classic microfacet BSDF.
+We have presented the Minimal Retroreflective Microfacet (MRM) based on a simple modification of the classic microfacet BSDF.
 We have shown that the model fulfills reciprocity and energy conservation requirements, making it a physically plausible microfacet model.

### 2025-12-23 MRM.tex
Correct author name.
Typo fix.

@@
         \and Laurent Belcour\\Intel
-        \and Frankie Liu\\NVIDIA
+        \and Francis Liu\\NVIDIA
         \and Jamie Portsmouth\\Autodesk
@@
 Figure~\ref{fig:data_fit} shows a fit of the retroreflective MRM model (in red) to the measured BRDF of retroreflective tape (in blue) from \cite{BelcourBackvector}.
-This is the "yellow tape" sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
+This is the ``yellow tape'' sample from their dataset, illuminated at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$ (in the plane of incidence).
 The MRM model provides a reasonably good match to the retroreflective peak using a roughness of $r = 0.23$, together with ``F82-tint'' Fresnel \cite{OpenPBR2024} with $F_0=0.19$ and $F_{82} = 0$.

### 2025-12-29 MRM.tex
Missing fullstop in equation.
Plot color improvements

@@
 \begin{equation}
-\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2}
+\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2} \ .
 \end{equation}
@@
   \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
-  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (red) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in \{45^\circ, 80^\circ\}$.
+  \caption{Comparison of the retroreflective MRM BRDF (green) to the regular GGX microfacet BRDF (blue) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in \{45^\circ, 80^\circ\}$.
 The dashed lines are the roughness $0.1$ case, and the solid lines roughness $0.666$.
-The horizontal green line at $1/\pi$ is the Lambert BRDF.}
+(The horizontal red line at $1/\pi$ is the Lambert BRDF).}
   \label{fig:brdf_plots}
@@
   \vspace{-0.5cm}
-  \caption{Measured BRDF of retroreflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
+  \caption{Measured BRDF of retroreflective tape (black dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (green dashed line).}
   \label{fig:data_fit}

### 2025-12-29 code/brdf_plotting.py

@@ -99,10 +99,10 @@ def plot_brdf_data():
             linestyle = 'solid'
             show_legend = True
 
-        axes[index].plot(data['theta_o'], data['Lambert'], label='Lambert' if show_legend else None, color='green',   linewidth=1.0, linestyle=linestyle)
+        axes[index].plot(data['theta_o'], data['Lambert'], label='Lambert' if show_legend else None, color='red',   linewidth=1.0, linestyle=linestyle)
         axes[index].plot(data['theta_o'], data['EON'],     label='EON'     if show_legend else None, color='orange', linewidth=1.0, linestyle=linestyle)
         axes[index].plot(data['theta_o'], data['GGX'],     label='GGX'     if show_legend else None, color='blue',   linewidth=1.0, linestyle=linestyle)
-        axes[index].plot(data['theta_o'], data['MRM'],     label='MRM'     if show_legend else None, color='red',  linewidth=1.0, linestyle=linestyle)
+        axes[index].plot(data['theta_o'], data['MRM'],     label='MRM'     if show_legend else None, color='green',  linewidth=1.0, linestyle=linestyle)
 
         # Set the title for each subplot
         axes[index].set_title(f"$\\theta_i: {theta_i}^\\circ$, $\\phi_o: {phi}^\\circ$", fontsize=14)

### 2025-12-29 code/brdf_plotting_with_data.py

@@ -110,7 +110,9 @@ for idx, (tL, data_points) in enumerate(sorted(brdf_dict_simple.items())):
 
     # Plot BRDF values vs theta_V (in degrees)
     axes[idx].clear()  # Clear previous plot
-    axes[idx].plot(theta_V_vals_deg, brdf_vals, 'o-', linewidth=1.5, markersize=4)
+    axes[idx].plot(theta_V_vals_deg, brdf_vals, 'o-', color='black', alpha=0.25, linewidth=1.5, markersize=0)
+    axes[idx].plot(theta_V_vals_deg, brdf_vals, 'o-', color='black', alpha=0.75, linewidth=0, markersize=3)
+
     axes[idx].set_xlabel('$\\theta_V$ (degrees)', fontsize=16)
     axes[idx].set_ylabel('BRDF value', fontsize=16)
     axes[idx].set_title(f'$\\theta_L = {tL_deg:.2f}^\\circ$', fontsize=16)
@@ -130,7 +132,7 @@ for idx, (tL, data_points) in enumerate(sorted(brdf_dict_simple.items())):
     axes[idx].axvline(x=tL_deg, color='grey', linestyle=':', linewidth=1.5, alpha=0.7)
     axes[idx].axvline(x=-tL_deg, color='grey', linestyle=':', linewidth=1.5, alpha=0.7)
 
-    axes[idx].plot(calc_brdf_dict[tL_deg][0], calc_brdf_dict[tL_deg][1], 'r--', label='Calculator MRM', linewidth=1.5)
+    axes[idx].plot(calc_brdf_dict[tL_deg][0], calc_brdf_dict[tL_deg][1], 'r--', color='green', label='Calculator MRM', linewidth=1.5)
 
 # Hide unused subplots
 for idx in range(len(brdf_dict_simple), len(axes)):

### 2025-12-29 MRM.tex
Wording fixes.

@@
 \citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface.
-Though here there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.
+Though here there is no direct physical basis for interpreting the back-vector as such a scattering mode, this can serve as an intuition for why the components of the regular microfacet model still work on changing the domain to the back-vector.
 
@@
 The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
-It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
+It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =\theta_i$, $\phi_o = \phi_i$).
 The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane as the light direction.
